---
title: "merge_exotic_status_authority_docs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The NEON working group has been using this csv file (taxonomy_temp10_revised.csv) as the authority for exotic status. However they have been overriding some of the exotic status assignations in taxonomy_temp10_revised.csv with the exotic status listed in this csv file: multStatusSpL48.xlsx

In this Rmarkdown I will be combining these files to create one file to use as the exotic status authority for the VegBank data. 

Loading packages

```{r}

library(readr)
library(readxl)
library(tidyverse)

```


```{r}

taxonomy_temp10_revised <- read_csv("/home/shares/neon-inv/raw_VegBank_data/taxonomy_temp10_revised.csv")

#Adding column with data origin so that when the datasets are merged we can keep track of which exotic status authorirty was used
taxonomy_temp10_revised <- taxonomy_temp10_revised %>% add_column(ExoticStatus_Origin = 'taxonomy_temp10_revised', .before = 'inv_L48') 

multStatusSpL48 <- read_excel("/home/shares/neon-inv/raw_VegBank_data/multStatusSpL48.xlsx")

#Adding column with data origin so that when the datasets are merged we can keep track of which exotic status authorirty was used
multStatusSpL48 <- multStatusSpL48 %>% add_column(ExoticStatus_Origin = 'multStatusSpL48', .before = 'FINAL DECISION (L48)') 

#Import USDA Plant List 
USDA_list_w_exotic_status <- read.delim2("/home/shares/neon-inv/raw_VegBank_data/USDA_Plant_List_020821.txt",  sep = ",")

#Adding column with data origin so that when the datasets are merged we can keep track of which exotic status authorirty was used
USDA_list_w_exotic_status <- USDA_list_w_exotic_status %>% add_column(ExoticStatus_Origin = 'USDA_PLANTLIST', .before = 'Native.Status') 

```

Read in USDA Plants list with species codes and only keep the native status for the L48

```{r}

#Only keeping L48 Native status by detecting string that start with L

USDA_list_w_exotic_status <- USDA_list_w_exotic_status %>%
  mutate(Native.Status = str_extract(`Native.Status`, "L[1-9]+..."))

#unique(USDA_list_w_exotic_status$Native.Status)

#Reducing the column to just include that status within the parenthesis ex. L48(N) = N
USDA_list_w_exotic_status$Native.Status <- substr(USDA_list_w_exotic_status$Native.Status, 5, 7)

USDA_list_w_exotic_status$Native.Status <- gsub("[()]", "", USDA_list_w_exotic_status$Native.Status)

USDA_list_w_exotic_status$Native.Status <- gsub('[[:punct:]]', "", USDA_list_w_exotic_status$Native.Status)

#Check to make sure this worked. I can see that the native status symbols that remain are "I", "N", "NI" and "W". I will have to ask the NEON team what they want to do about the 'Waif' status individuals if they are present. There are only 16 waifs and 19 NI's so its mostly I and N species. 
unique(USDA_list_w_exotic_status$Native.Status)

summary(USDA_list_w_exotic_status$Native.Status)

#Cleaning up the table by removing some extra columns we don't need and renaming others
USDA_plant_list_ALL <- USDA_list_w_exotic_status %>%
  select(-"Symbol") %>%
 rename("bestname" = "Scientific.Name", "SpCode" = 'Accepted.Symbol', "USDA_Exotic_Status" = "Native.Status", "USDA_Growth_Form" = "Growth.Habit", "USDA_Duration" = "Duration") 



```

Reducing the USDA growth form description down to one word

```{r}

USDA_plant_list_ALL$USDA_Growth_Form <- gsub('[[:punct:]]', " ", USDA_plant_list_ALL$USDA_Growth_Form)

USDA_plant_list_ALL$USDA_Growth_Form <- stringr::word(USDA_plant_list_ALL$USDA_Growth_Form, 1)

unique(USDA_plant_list_ALL$USDA_Growth_Form)

```

Selecting only the columns I need from the csv file and excel file provided by the NEON working group.

During our meeting I asked Lais if the 'inv_L48' column was the best one to use for exotic status and she said yes. I did not ask which column I should use for 'bestname'. There is a bestname column already in the table BUT when I use the bestname = bestname I end up with duplicates. To see what I mean by this look at this:

```{r}

taxonomy_example <- read_csv("/home/shares/neon-inv/raw_VegBank_data/taxonomy_temp10_revised.csv")

#Look at rows 7 and 8 of this table. They have the exact same 'bestname' but different 'Scientific.Name's and different SpCodes. It would make sense to me to use the bestname column because it already has the name of the column I need but when I do I end up with duplicates. 

```


```{r}

taxonomy_temp10_revised <- taxonomy_temp10_revised %>% select('Symbol', 'inv_L48', 'GrowthForm', 'Scientific.Name', 'ExoticStatus_Origin') %>% rename('SpCode' = 'Symbol', 'NEON_GrowthForm' = 'GrowthForm', 'bestname' = 'Scientific.Name')

multStatusSpL48 <- multStatusSpL48 %>% select('SpCode', 'FINAL DECISION (L48)', 'ExoticStatus_Origin') %>% rename('Exception_Exotic_Status' = 'FINAL DECISION (L48)')

exotic_status1 <- dplyr::left_join(taxonomy_temp10_revised, multStatusSpL48, by = 'SpCode')

exotic_status2 <- exotic_status1 %>%
  mutate('NEW_ExoticStatus' = ifelse(!is.na(Exception_Exotic_Status), Exception_Exotic_Status, inv_L48))

exotic_status3 <- exotic_status2 %>%
  mutate('ExoticStatus_Origin' = ifelse(!is.na(ExoticStatus_Origin.x), ExoticStatus_Origin.x, ExoticStatus_Origin.y))

KPEACH_ExoticStatus_authority1 <- exotic_status3 %>% select('SpCode', 'NEW_ExoticStatus', 'NEON_GrowthForm', 'bestname', 'ExoticStatus_Origin')

KPEACH_ExoticStatus_authority1  <- KPEACH_ExoticStatus_authority1[!duplicated(KPEACH_ExoticStatus_authority1$bestname), ]

```

For any new species that are not already listed in the documents provided by NEON (taxonomy_temp10_revised.csv and multStatusSpL48.xlsx) I am using the USDA exotic status.

```{r}

KPEACH_ExoticStatus_authority2 <- full_join(KPEACH_ExoticStatus_authority1, USDA_plant_list_ALL, by = 'bestname')

KPEACH_ExoticStatus_authority3 <- KPEACH_ExoticStatus_authority2 %>%
  mutate('NEW_ExoticStatus' = ifelse(!is.na(NEW_ExoticStatus), NEW_ExoticStatus, USDA_Exotic_Status))

KPEACH_ExoticStatus_authority4 <- KPEACH_ExoticStatus_authority3 %>%
  mutate('ExoticStatus_Origin' = ifelse(!is.na(ExoticStatus_Origin.x), ExoticStatus_Origin.x, ExoticStatus_Origin.y))

KPEACH_ExoticStatus_authority <- KPEACH_ExoticStatus_authority4 %>% select(-'SpCode.y', -'ExoticStatus_Origin.x', -'ExoticStatus_Origin.y') %>% rename('SpCode' = 'SpCode.x')

KPEACH_ExoticStatus_authority  <- KPEACH_ExoticStatus_authority[!duplicated(KPEACH_ExoticStatus_authority$bestname), ]

```

There are a couple species that appear on the USDA website with valid SpCode but were not exported with the list. Manually entering these here

```{r}

#https://plants.usda.gov/core/profile?symbol=PALY3
KPEACH_ExoticStatus_authority <- KPEACH_ExoticStatus_authority %>% 
add_row(bestname = 'Pallavicinia lyellii', SpCode = 'PALY3', USDA_Exotic_Status = "N", USDA_Growth_Form = "Nonvascular")

```


Saving output files locally and in shared folder on aurora server
```{r}

#As csv on server in shared folder
write.csv(KPEACH_ExoticStatus_authority,"/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority.csv", row.names = FALSE)

#save locally
write.csv(KPEACH_ExoticStatus_authority,"KPEACH_ExoticStatus_authority.csv", row.names = FALSE)

```

