---
title: "VegBank_FL_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Florida VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}


VegBank_plot_data_FL <- read_excel("CVS-FLplots.xlsx", sheet = "plot data")
VegBank_plot_species_list_FL <- read_excel("CVS-FLplots.xlsx",  sheet = "plot species list")
VegBank_taxon_importance_FL <- read_excel(here::here("CVS-FLplots.xlsx"),  sheet = "taxon importance")
VegBank_soil_avg_by_horizon_FL <- read_excel(here::here("CVS-FLplots.xlsx"), sheet = "soil data avgd by horizon")
VegBank_disturbances_FL <- read_excel(here::here("CVS-FLplots.xlsx"), sheet = "disturbances")
VegBank_species_richness_FL <-  read_excel(here::here("CVS-FLplots.xlsx"), sheet = "plot species richness")

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}


VegBank_plot_data_FL2 <- VegBank_plot_data_FL %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'UniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
  filter('Sampling Level' > 4)

VegBank_plot_species_list_FL2 <- VegBank_plot_species_list_FL %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic', 'taxon_usda_growthForm', 'taxon_usda_code', 'taxon_woody') %>%
  rename( 'UniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_FL <- full_join(VegBank_plot_data_FL2, VegBank_plot_species_list_FL2, by = 'UniqueID')

#Reducing sampling date to just the year to match NEON data
VegBank_FL <- VegBank_FL %>% 
  mutate('Year' = substr(VegBank_FL$Year, 1, 4))

#Adding a site column
VegBank_FL <- VegBank_FL %>% add_column(Dataset = 'VEGBANK', .before = "UniqueID")  %>% add_column(Zone = 'L48', .before = "UniqueID")

#Making exotic status compatible to NEON data
VegBank_FL$ExoticStatus <- gsub("native", "N", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("NATIVE", "N", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("exotic", "NI", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("EXOTIC", "NI", VegBank_FL$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_FL$Woodiness <- gsub("Y", "Woody", VegBank_FL$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_FL <- VegBank_FL  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))

#This just replaces all NA's, any row that was not explicitly labeled as 'Woody" as Herbaceous which could be wrong so I did not run it
#VegBank$Woodiness<-  replace_na(VegBank$Woodiness, "Herbaceous")

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_FL$bestname <- gsub("subsp.", "ssp.", VegBank_FL$bestname)
VegBank_FL$bestname <- gsub("subspecies", "ssp.", VegBank_FL$bestname)
VegBank_FL$bestname <- gsub("ssp..", "ssp.", VegBank_FL$bestname)

#strip white space
VegBank_FL$bestname  <- trimws(VegBank_FL$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_FL <- VegBank_FL %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_FL$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_FL$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrides
VegBank_multiple_species_FL$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_FL$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_FL ,"multiple_sp_FL_KPEACH.csv", row.names = FALSE)

```

Removing these rows from the main data set until I can find a solution
```{r}

VegBank_FL <- VegBank_FL %>%
    filter(!str_detect(bestname,  ".\\[|\\]."))


```

Read in USDA Plants list with species codes

Not sure why it would not let me download all of them at once?

```{r}


USDA_plant_list_ALL <- read.delim2(here::here("USDA_all_with_syn.txt"),  sep = ",")

USDA_plant_list_ALL <- USDA_plant_list_ALL %>%
  rename("bestname" = "Scientific.Name", "SpCode" = 'Accepted.Symbol') 

```


Manually fixing some species names

```{r}

VegBank_FL$bestname <- gsub("Dichanthelium tenue", "Dichanthelium dichotomum var. tenue", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia anceps ssp.rhizomata", "Panicum anceps", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Vaccinium stamineum var. caesium", "Vaccinium stamineum", VegBank_FL$bestname)


VegBank_FL$bestname <- gsub("Coleataenia longifolia ssp.longifolia", "Coleataenia longifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Galactia volubilis var. volubilis", "Galactia volubilis", VegBank_FL$bestname)
	

VegBank_FL$bestname <- gsub("Triplasis purpurea var. purpurea", "Triplasis purpurea", VegBank_FL$bestname)



```




Table of species with no match in USDA table

```{r}

#1,358 rows with no match
no_match <- dplyr::anti_join(VegBank_FL,USDA_plant_list_ALL, by="bestname")

#Reducing all rows of the no match file to just the genus so that USDA can find matches
no_match$bestname <- stringr::word(no_match$bestname, 1)

#Save as csv for Jeff to inspect
write.csv(no_match,"no_match_FL_KPEACH.csv", row.names = FALSE)

```



Adding species codes to rows with matches

```{r}

#24,042
VegBank_with_SpCode_FL1 <- dplyr::inner_join(VegBank_FL, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_FL2 <- dplyr::inner_join(no_match, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_FL <- merge(VegBank_with_SpCode_FL1, VegBank_with_SpCode_FL2, all = TRUE)


VegBank_with_SpCode_FL <- VegBank_with_SpCode_FL %>%
  select(-taxon_usda_growthForm, -SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Now lets see how many no match rows are left
#0! Yay!
no_match2 <- dplyr::anti_join(VegBank_with_SpCode_FL,USDA_plant_list_ALL, by="bestname")

```

```{r}

#As csv
write.csv(VegBank_with_SpCode_FL,"VegBank_FL_KPEACH.csv", row.names = FALSE)

```