---
title: "VegBank_GA_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Georgia VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_GA <- read_excel(file.path(data_raw, "CVS-GAplots.xlsx"), sheet = "plot data", guess_max = min(4000, n_max = NULL))
VegBank_plot_species_list_GA <- read_excel(file.path(data_raw, "CVS-GAplots.xlsx"),  sheet = "plot species list", guess_max = min(4000, n_max = NULL))

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")

```

Extra cleanup

```{r}

VegBank_plot_data_GA$`Observation Start Date` <- gsub("UTC", " ", VegBank_plot_data_GA$`Observation Start Date`)

#unique(VegBank_plot_data_GA$`Observation Start Date`)

VegBank_plot_data_GA <- VegBank_plot_data_GA %>% drop_na(`Observation Start Date`)

VegBank_plot_data_GA <- VegBank_plot_data_GA %>%
   filter(str_detect(`Taxon Observation Area`, "\\d")) 

VegBank_plot_species_list_GA <- VegBank_plot_species_list_GA %>% drop_na(currentTaxonNameWithSp)

VegBank_plot_species_list_GA$cover <- as.numeric(VegBank_plot_species_list_GA$cover)

unique(VegBank_plot_species_list_GA$cover)

VegBank_plot_species_list_GA <- VegBank_plot_species_list_GA %>% drop_na(cover)


```

There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the data before I even really touch it.

1 Duplicate row is in the data before I do anything to it.

```{r}

duplicates1 <- VegBank_plot_species_list_GA %>% group_by(authorObsCode, currentTaxonNameWithSp) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) %>%  filter(currentTaxonNameWithSp != "{Unknown}") 

```


Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}

VegBank_plot_data_GA$`Sampling Level` <- as.numeric(VegBank_plot_data_GA$`Sampling Level`)


VegBank_plot_data_GA2 <- VegBank_plot_data_GA %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)  

VegBank_plot_species_list_GA2 <- VegBank_plot_species_list_GA %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_GA <- left_join(VegBank_plot_data_GA2, VegBank_plot_species_list_GA2, by = 'VegBankUniqueID')

#Check to make sure only Sampling Levels 4 and 5 are included
unique(VegBank_GA$`Sampling Level`)


```
```{r}

#Reducing sampling date to just the year to match NEON data
VegBank_GA <- VegBank_GA %>% 
  mutate('Year' = substr(VegBank_GA$Year, 1, 4))

#Adding a site column
VegBank_GA <- VegBank_GA %>% add_column(Dataset = 'VEGBANK_GA', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

#Making exotic status compatible to NEON data
VegBank_GA$ExoticStatus <- gsub("native", "N", VegBank_GA$ExoticStatus)
VegBank_GA$ExoticStatus <- gsub("NATIVE", "N", VegBank_GA$ExoticStatus)
VegBank_GA$ExoticStatus <- gsub("exotic", "I", VegBank_GA$ExoticStatus)
VegBank_GA$ExoticStatus <- gsub("EXOTIC", "I", VegBank_GA$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_GA$Woodiness <- gsub("Y", "Woody", VegBank_GA$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_GA <- VegBank_GA  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))


#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_GA$bestname <- gsub("subsp.", "ssp.", VegBank_GA$bestname, fixed = TRUE)
VegBank_GA$bestname <- gsub("subspecies", "ssp.", VegBank_GA$bestname, fixed = TRUE)
VegBank_GA$bestname <- gsub("ssp..", "ssp.", VegBank_GA$bestname, fixed = TRUE)

VegBank_GA$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_GA$bestname, ignore.case = FALSE)

#strip white space
VegBank_GA$bestname  <- trimws(VegBank_GA$bestname, which = c("right"))





```



Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_GA <- VegBank_GA %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_GA$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_GA$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_GA$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_GA$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_GA ,"/home/shares/neon-inv/output_files/multiple_sp_GA_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows from the main data set until I can find a solution. 
```{r}

VegBank_GA$bestname  <- str_replace_all(VegBank_GA$bestname , "\\*|\\[|\\]", "")

```

Manually Fixing some species

```{r}

VegBank_GA$bestname <- gsub("Liatris sp.", "Liatris", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Erianthus sp.", "Erianthus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Viola pedata var. pedata", "Viola pedata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Hamamelis virginiana var. virginiana", "Hamamelis virginiana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Betula lenta var. lenta", "Betula lenta", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum ssp. racemosum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Magnolia acuminata var. acuminata", "Magnolia acuminata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Pteridium latiusculum var. latiusculum", "Pteridium aquilinum var. latiusculum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Muhlenbergia sp.", "Muhlenbergia", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Viola sp.", "Viola", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Nabalus", "Prenanthes", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Helianthus sp.", "Helianthus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Lespedeza sp.", "Lespedeza", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Luzula sp.", "Luzula", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Hypericum sp.", "Hypericum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Solidago sp.", "Solidago", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Stachys sp.", "Stachys", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Erianthus", "Saccharum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Oenothera sp.", "Oenothera", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Rubus sp.", "Rubus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Quercus sp.", "Quercus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Agrostis sp.", "Agrostis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bidens sp.", "Bidens", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Dichanthelium sp.", "Dichanthelium", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Erigeron sp.", "Erigeron", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Dichanthelium tenue", "Dichanthelium dichotomum var. tenue", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Symphyotrichum concolor var. concolor", "Symphyotrichum concolor", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Clitoria mariana var. mariana", "Clitoria mariana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Polygonum americanum", "Polygonella americana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Seymeria pectinata ssp.pectinata", "Seymeria pectinata ssp. pectinata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Ionactis linariifolia", "Ionactis linariifolius", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Vaccinium sp.", "Vaccinium", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Cyperus lupulinus var. lupulinus", "Cyperus lupulinus ssp. lupulinus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Agalinis sp.", "Agalinis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Crataegus sp.", "Crataegus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Lechea sp.", "Lechea", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Dichantheliumaerocarpon", "Dichanthelium sphaerocarpon", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Chasmanthium sessiliflorum var. sessiliflorum", "Chasmanthium sessiliflorum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Opuntia mesacantha ssp. mesacantha", "Opuntia humifusa", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Crataegusthulata", "Crataegus spathulata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bigelowia nudata var. nudata", "Bigelowia nudata ssp. nudata", VegBank_GA$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=1010
VegBank_GA$bestname <- gsub("Solidago virgata", "Solidago sempervirens var. mexicana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Orbexilum lupinellus", "Orbexilum lupinellum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Saccharum giganteus", "Saccharum giganteum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Andropogon gerardi", "Andropogon gerardii", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Orbexilum psoralioides", "Orbexilum pedunculatum var. psoralioides", VegBank_GA$bestname)

#https://swbiodiversity.org/seinet/taxa/index.php?taxon=sherardia%20arvensis
VegBank_GA$bestname <- gsub("Galium sherardia", "Sherardia arvensis", VegBank_GA$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3790
VegBank_GA$bestname <- gsub("Polygonum pinicola", "Polygonella gracilis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Ludwigia linearis var. linearis", "Ludwigia linearis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Cartrema americanum", "Osmanthus americanus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Sarracenia minor var. minor", "Sarracenia minor", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Dichanthelium chamaelonche ssp. chamaelonche", "Dichanthelium dichotomum var. breve", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Lithospermum virginianum", "Onosmodium virginianum", VegBank_GA$bestname)

#https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi?id=57934
VegBank_GA$bestname <- gsub("Sibbaldia retusa", "Sibbaldiopsis tridentata", VegBank_GA$bestname)

#https://www.marylandbiodiversity.com/viewSpecies.php?species=4078
VegBank_GA$bestname <- gsub("Hymenachne hemitomon", "Panicum hemitomon", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Rhynchospora cephalantha var. cephalantha", "Rhynchospora cephalantha", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bryodesma rupestre", "Selaginella rupestris", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bryodesma tortipilum", "Selaginella tortipila", VegBank_GA$bestname)

#https://plants.ces.ncsu.edu/plants/hemionitis-lanosa/
VegBank_GA$bestname <- gsub("Myriopteris lanosa", "Cheilanthes lanosa", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bryodesma arenicola", "Selaginella arenicola", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Zephyranthes atamasco", "Zephyranthes atamasca", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Glyceria striata var. striata", "Glyceria striata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Smallanthus uvedalia", "Smallanthus uvedalius", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Styrax americanus var. americanus", "Styrax americanus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Aureolaria levigata", "Aureolaria laevigata", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Prenanthes altissimus", "Prenanthes altissima", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Bryodesma tortipilum", "Selaginella tortipila", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Coleataenia anceps ssp. rhizomata", "Panicum anceps", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Coleataenia longifolia ssp. longifolia", "Panicum rigidulum var. pubescens", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Saccharum coarctatus", "Saccharum coarctatum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Vaccinium stamineum var. caesium", "Vaccinium simulatum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Vaccinium stamineum var. stamineum", "Vaccinium stamineum", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Morella pumila", "Morella cerifera", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Andropogon dealbatus", "Andropogon capillipes", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Tiedemannia filiformis ssp. filiformis", "Oxypolis filiformis", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Osmunda claytoniana var. claytoniana", "Osmunda claytoniana", VegBank_GA$bestname)

#There is no Dioscorea species really close to polystachya but this species has the same common name
VegBank_GA$bestname <- gsub("Dioscorea polystachya", "Dioscorea oppositifolia", VegBank_GA$bestname)

#https://swbiodiversity.org/seinet/taxa/index.php?taxon=221552&clid=4026
VegBank_GA$bestname <- gsub("Mononeuria glabra", "Minuartia glabra", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Myriopteris tomentosa", "Cheilanthes tomentosa", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Poa trivialis ssp. trivialis", "Poa trivialis", VegBank_GA$bestname)

#https://plants.ces.ncsu.edu/plants/homalosorus-pycnocarpos/
VegBank_GA$bestname <- gsub("Homalosorus pycnocarpos", "Diplazium pycnocarpon", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Cotinus obovata", "Cotinus obovatus", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Solidagoacelata", "Solidago sphacelata", VegBank_GA$bestname)

#http://www.namethatplant.net/plantdetail.shtml?plant=988
VegBank_GA$bestname <- gsub("Opuntia mesacantha", "Opuntia humifusa", VegBank_GA$bestname)

VegBank_GA$bestname <- gsub("Liatriscata", "Liatris spicata", VegBank_GA$bestname)


```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_GA, KPEACH_ExoticStatus_authority, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "{Unknown}") %>%
 filter(!str_detect(bestname, "\\d"))


```

```{r}

no_match <- unique(no_match)

```


Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_GA", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```

Adding species codes to rows with matches

```{r}

VegBank_GA <- unique(VegBank_GA)

VegBank_with_SpCode_GA1 <- dplyr::inner_join(VegBank_GA, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_GA1 <- VegBank_with_SpCode_GA1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_GA <- merge(VegBank_with_SpCode_GA1, no_match, all = TRUE)

VegBank_with_SpCode_GA <- VegBank_with_SpCode_GA %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')

```


I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 


```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_GA$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_GA$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_GA$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_GA$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_GA <- VegBank_with_SpCode_GA %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))



```

Renaming USDA_Duration column to just 'Duration' 

```{r}

VegBank_GA5 <- VegBank_with_SpCode_GA %>% rename('Duration' = 'USDA_Duration') %>% select(-Synonym.Symbol)


```

Looking for source of duplicates

```{r}

#4 duplicates. Identical except for cover value

duplicates <- VegBank_GA5 %>% group_by(VegBankUniqueID, SpCode) %>% summarise(n_obs = n()) %>% filter(n_obs > 1)

```


Exporting full data table

```{r}

#As csv
write.csv(VegBank_GA5,"/home/shares/neon-inv/output_files/VegBank_GA_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

VegBank_GA6 <- VegBank_GA5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_GA6 <- VegBank_GA6 %>% drop_na(Lat)

#As csv
write.csv(VegBank_GA6,"/home/shares/neon-inv/output_files/VegBank_GA_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_GA6,"VegBank_GA_KPEACH_reduced.csv", row.names = FALSE)

```