---
title: "VegBank_MS_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Mississippi VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}


# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_MS <- read_excel(file.path(data_raw, "CVS-MSplots.xlsx"), sheet = "plot data")
VegBank_plot_species_list_MS <- read_excel(file.path(data_raw, "CVS-MSplots.xlsx"),  sheet = "plot species list")

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")


```

Extra cleanup

```{r}

VegBank_plot_data_MS$`Observation Start Date` <- gsub("UTC", " ", VegBank_plot_data_MS$`Observation Start Date`)

#unique(VegBank_plot_data_MS$`Observation Start Date`)

VegBank_plot_data_MS <- VegBank_plot_data_MS %>% drop_na(`Observation Start Date`)

VegBank_plot_data_MS <- VegBank_plot_data_MS %>%
   filter(str_detect(`Taxon Observation Area`, "\\d")) 

VegBank_plot_species_list_MS <- VegBank_plot_species_list_MS %>% drop_na(currentTaxonNameWithSp)

VegBank_plot_species_list_MS$cover <- as.numeric(VegBank_plot_species_list_MS$cover)

unique(VegBank_plot_species_list_MS$cover)

VegBank_plot_species_list_MS <- VegBank_plot_species_list_MS %>% drop_na(cover)


```
There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the data before I even really touch it.

0 Duplicate rows are in the data before I do anything to it.

```{r}

duplicates1 <- VegBank_plot_species_list_MS %>% group_by(authorObsCode, currentTaxonNameWithSp) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) %>%  filter(currentTaxonNameWithSp != "{Unknown}") 

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}

VegBank_plot_data_MS$`Sampling Level` <- as.numeric(VegBank_plot_data_MS$`Sampling Level`)

VegBank_plot_data_MS2 <- VegBank_plot_data_MS %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)  

VegBank_plot_species_list_MS2 <- VegBank_plot_species_list_MS %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_MS <- left_join(VegBank_plot_data_MS2, VegBank_plot_species_list_MS2, by = 'VegBankUniqueID')

#Check to make sure only Sampling Levels 4 and 5 are included
unique(VegBank_MS$`Sampling Level`)



```


```{r}

#Reducing sampling date to just the year to match NEON data
VegBank_MS <- VegBank_MS %>% 
  mutate('Year' = substr(VegBank_MS$Year, 1, 4))

#Adding a site column
VegBank_MS <- VegBank_MS %>% add_column(Dataset = 'VEGBANK_MS', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

#Making exotic status compatible to NEON data
VegBank_MS$ExoticStatus <- gsub("native", "N", VegBank_MS$ExoticStatus)
VegBank_MS$ExoticStatus <- gsub("NATIVE", "N", VegBank_MS$ExoticStatus)
VegBank_MS$ExoticStatus <- gsub("exotic", "I", VegBank_MS$ExoticStatus)
VegBank_MS$ExoticStatus <- gsub("EXOTIC", "I", VegBank_MS$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_MS$Woodiness <- gsub("Y", "Woody", VegBank_MS$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_MS <- VegBank_MS  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))


#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_MS$bestname <- gsub("subsp.", "ssp.", VegBank_MS$bestname)
VegBank_MS$bestname <- gsub("subspecies", "ssp.", VegBank_MS$bestname)
VegBank_MS$bestname <- gsub("ssp..", "ssp.", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_MS$bestname, ignore.case = FALSE)

#strip white space
VegBank_MS$bestname  <- trimws(VegBank_MS$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_MS <- VegBank_MS %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_MS$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_MS$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrides
VegBank_multiple_species_MS$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_MS$bestname)

#Save as csv for Jeff to inspect
#write.csv(VegBank_multiple_species_MS ,"multiple_sp_MS_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows from the main data set until I can find a solution. 
```{r}

VegBank_MS$bestname  <- str_replace_all(VegBank_MS$bestname , "\\*|\\[|\\]", "")

```

Manually fixing some species names

```{r}


VegBank_MS$bestname <- gsub("Cyperus lupulinus var. lupulinus", "Cyperus lupulinus ssp. lupulinus", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Dichantheliumaerocarpon", "Dichanthelium sphaerocarpon", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_MS$bestname)

VegBank_MS$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum ssp. racemosum", VegBank_MS$bestname)




```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_MS, KPEACH_ExoticStatus_authority, by="bestname")

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "{Unknown}") %>%
 filter(!str_detect(bestname, "\\d"))


```

```{r}

no_match <- unique(no_match)

```


Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_MS", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```

Adding species codes to rows with matches

```{r}

VegBank_MS <- unique(VegBank_MS)

VegBank_with_SpCode_MS1 <- dplyr::inner_join(VegBank_MS, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_MS1 <- VegBank_with_SpCode_MS1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_MS <- merge(VegBank_with_SpCode_MS1, no_match, all = TRUE)

VegBank_with_SpCode_MS <- VegBank_with_SpCode_MS %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 


```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_MS$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_MS$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_MS$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_MS$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_MS <- VegBank_with_SpCode_MS %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))


```

Renaming USDA_Duration column to just 'Duration' per Lais' request

```{r}

VegBank_MS5 <- VegBank_with_SpCode_MS %>% rename('Duration' = 'USDA_Duration') %>% select(-Synonym.Symbol)


```

Looking for source of duplicates

```{r}

#4 duplicates. Identical except for cover value

duplicates <- VegBank_MS5 %>% group_by(VegBankUniqueID, SpCode) %>% summarise(n_obs = n()) %>% filter(n_obs > 1)

```


Exporting full data table

```{r}

#As csv
write.csv(VegBank_MS5,"/home/shares/neon-inv/output_files/VegBank_MS_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

VegBank_MS6 <- VegBank_MS5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_MS6 <- VegBank_MS6 %>% drop_na(Lat)

#As csv
write.csv(VegBank_MS6,"/home/shares/neon-inv/output_files/VegBank_MS_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_MS6,"VegBank_MS_KPEACH_reduced.csv", row.names = FALSE)

```