---
title: "VegBank_Project129_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Project 129 VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}


# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_Proj129 <- read_excel(file.path(data_raw, "CVS-Project129_new.xlsx"), sheet = "plot data")
VegBank_plot_species_list_Proj129 <- read_excel(file.path(data_raw, "CVS-Project129_new.xlsx"),  sheet = "plot species list")

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority.csv")

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}


VegBank_plot_data_Proj129$`Sampling Level` <- as.numeric(VegBank_plot_data_Proj129$`Sampling Level`)

VegBank_plot_data_Proj1292 <- VegBank_plot_data_Proj129 %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)  

VegBank_plot_species_list_Proj1292 <- VegBank_plot_species_list_Proj129 %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_Proj129 <- left_join(VegBank_plot_data_Proj1292, VegBank_plot_species_list_Proj1292, by = 'VegBankUniqueID')

#Reducing sampling date to just the year to match NEON data
VegBank_Proj129 <- VegBank_Proj129 %>% 
  mutate('Year' = substr(VegBank_Proj129$Year, 1, 4))

#Adding a site column
VegBank_Proj129 <- VegBank_Proj129 %>% add_column(Dataset = 'VEGBANK', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

#Making exotic status compatible to NEON data
VegBank_Proj129$ExoticStatus <- gsub("native", "N", VegBank_Proj129$ExoticStatus)
VegBank_Proj129$ExoticStatus <- gsub("NATIVE", "N", VegBank_Proj129$ExoticStatus)
VegBank_Proj129$ExoticStatus <- gsub("exotic", "I", VegBank_Proj129$ExoticStatus)
VegBank_Proj129$ExoticStatus <- gsub("EXOTIC", "I", VegBank_Proj129$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_Proj129$Woodiness <- gsub("Y", "Woody", VegBank_Proj129$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_Proj129 <- VegBank_Proj129  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))


#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_Proj129$bestname <- gsub("subsp.", "ssp.", VegBank_Proj129$bestname)
VegBank_Proj129$bestname <- gsub("subspecies", "ssp.", VegBank_Proj129$bestname)
VegBank_Proj129$bestname <- gsub("ssp..", "ssp.", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("\\bSP\\.\\b", " ", VegBank_Proj129$bestname)

#strip white space
VegBank_Proj129$bestname  <- trimws(VegBank_Proj129$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_Proj129 <- VegBank_Proj129 %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_Proj129$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_Proj129$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrides
VegBank_multiple_species_Proj129$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_Proj129$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_Proj129 ,"multiple_sp_Proj129_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows from the main data set 
```{r}

VegBank_Proj129$bestname  <- str_replace_all(VegBank_Proj129$bestname , "\\*|\\[|\\]", "")

VegBank_Proj129$bestname  <- str_replace_all(VegBank_Proj129$bestname , "\\*|\\{|\\}", "")

```

Manually fixing some species names

```{r}


VegBank_Proj129$bestname <- gsub("Viola pedata var. pedata", "Viola pedata", VegBank_Proj129$bestname)


VegBank_Proj129$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Cartrema americanum", "Osmanthus americanus", VegBank_Proj129$bestname)

#http://www.namethatplant.net/plantdetail.shtml?plant=988
VegBank_Proj129$bestname <- gsub("Opuntia mesacantha", "Opuntia humifusa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Hamamelis virginiana var. virginiana", "Hamamelis virginiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Arthraxon hispidus var. hispidus", "Arthraxon hispidus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Magnolia virginiana var. virginiana", "Magnolia virginiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Iva frutescens var. frutescens", "Iva frutescens ssp. frutescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Gonolobus suberosus var. suberosus", "Matelea gonocarpos", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Paronychia baldwinii ssp.riparia", "Paronychia baldwinii ssp. riparia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Aureolaria levigata", "Aureolaria laevigata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Paronychia baldwinii ssp.riparia", "Paronychia baldwinii ssp. riparia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Hypericum mutilum var. mutilum", "Hypericum mutilum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Clitoria mariana var. mariana", "Clitoria mariana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium portoricense ssp.patulum", "Dichanthelium sabulorum var. thinium", VegBank_Proj129$bestname)


VegBank_Proj129$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Arthraxon hispidus var. hispidus", "Arthraxon hispidus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Phoradendron leucarpum ssp.leucarpum", "Phoradendron leucarpum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Chamaecyparis thyoides var. thyoides", "Chamaecyparis thyoides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Lepidium virginicum ssp.virginicum", "Lepidium virginicum var. virginicum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Arenaria lanuginosa var. lanuginosa", "Arenaria lanuginosa ssp. lanuginosa var. lanuginosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Verbascum thapsus ssp.thapsus", "Verbascum thapsus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Veronica peregrina var. peregrina", "Veronica peregrina ssp. peregrina", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Festuca rubra ssp.rubra", "Festuca rubra ssp. rubra", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Paronychia baldwinii ssp.baldwinii", "Paronychia baldwinii ssp. baldwinii", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Triplasis purpurea var. purpurea", "Triplasis purpurea", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Galactia volubilis var. volubilis", "Galactia volubilis", VegBank_Proj129$bestname)

#https://www.itis.gov/servlet/SingleRpt/SingleRpt?search_topic=TSN&search_value=893824#null
VegBank_Proj129$bestname <- gsub("Seutera angustifolia", "Cynanchum angustifolium", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Galium obtusum var. obtusum", "Galium obtusum ssp. obtusum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Opuntia humifusa ssp.mesacantha", "Opuntia humifusa", VegBank_Proj129$bestname)

#https://www.marylandbiodiversity.com/viewSpecies.php?species=4078
VegBank_Proj129$bestname <- gsub("Hymenachne hemitomon", "Panicum hemitomon", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Nymphaea odorata ssp.odorata", "Nymphaea odorata ssp. odorata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Tripsacum dactyloides var. dactyloides", "Tripsacum dactyloides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Hedera helix var. helix", "Hedera helix", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coleataenia longifolia ssp.longifolia", "Panicum rigidulum var. pubescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Pontederia cordata var. cordata", "Pontederia cordata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Oenothera fruticosa var. fruticosa", "Oenothera fruticosa ssp. fruticosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Sagittaria latifolia var. latifolia", "Sagittaria latifolia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_Proj129$bestname)

#https://www.itis.gov/servlet/SingleRpt/SingleRpt;jsessionid=301BE5D420CAB47EF9DE4C2B0476FA5E?search_topic=TSN&search_value=1061799#null
VegBank_Proj129$bestname <- gsub("Dichanthelium neuranthum", "Dichanthelium aciculare", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Galium tinctorium var. tinctorium", "Galium tinctorium", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Gaillardia pulchella var. drummondii", "Gaillardia pulchella", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/pourthiaea-villosa/
VegBank_Proj129$bestname <- gsub("Pourthiaea villosa", "Photinia villosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Erianthus", "Saccharum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Saccharum giganteus", "Saccharum giganteum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_Proj129$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3967
VegBank_Proj129$bestname <- gsub("Solidago virgata", "Solidago stricta", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Rumex crispus ssp.crispus", "Rumex crispus ssp. crispus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Glyceria striata var. striata", "Glyceria striata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Eragrostis secundiflora ssp.oxylepis", "Eragrostis secundiflora ssp. oxylepis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Kali soda", "Salsola kali ssp. kali", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Styrax americanus var. americanus", "Styrax americanus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium dichotomum var. ramulosum", "Dichanthelium dichotomum var. dichotomum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Morella pumila", "Morella cerifera", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Symphyotrichum concolor var. concolor", "Symphyotrichum concolor", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Phlox nivalis var. nivalis", "Phlox nivalis ssp. nivalis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Vernonia angustifolia var. angustifolia", "Vernonia angustifolia ssp. angustifolia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Bigelowia nudata var. nudata", "Bigelowia nudata ssp. nudata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Bryodesma rupestre", "Selaginella rupestris", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Bryodesma tortipilum", "Selaginella tortipila", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/hemionitis-lanosa/
VegBank_Proj129$bestname <- gsub("Myriopteris lanosa", "Cheilanthes lanosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Bryodesma arenicola", "Selaginella arenicola", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Zephyranthes atamasco", "Zephyranthes atamasca", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Orbexilum psoralioides", "Orbexilum pedunculatum var. psoralioides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ionactis linariifolia", "Ionactis linariifolius", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Chamaecrista nictitans var. nictitans", "Chamaecrista nictitans ssp. nictitans", VegBank_Proj129$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3080
VegBank_Proj129$bestname <- gsub("Dichanthelium webberianum", "Dichanthelium sabulorum var. thinium", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium tenue", "Dichanthelium dichotomum var. tenue", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Andropogon gerardi", "Andropogon gerardii", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Comandra umbellata var. umbellata", "Comandra umbellata ssp. umbellata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Asclepias tuberosa var. tuberosa", "Asclepias tuberosa ssp. tuberosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Sarracenia rubra ssp.rubra", "Sarracenia rubra ssp. rubra", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Orbexilum lupinellus", "Orbexilum lupinellum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium chamaelonche ssp.chamaelonche", "Dichanthelium dichotomum var. breve", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Bulbostylis capillaris ssp.capillaris", "Bulbostylis capillaris ssp. capillaris", VegBank_Proj129$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3399
VegBank_Proj129$bestname <- gsub("Lithospermum virginianum", "Onosmodium virginianum", VegBank_Proj129$bestname)

#https://swbiodiversity.org/seinet/taxa/index.php?taxon=100793
VegBank_Proj129$bestname <- gsub("Bryodesma acanthonota", "Selaginella acanthonota", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Sarracenia minor var. minor", "Sarracenia minor", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Rhynchospora glomerata var. glomerata", "Rhynchospora glomerata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum ssp. racemosum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Populus deltoides ssp.deltoides", "Populus deltoides ssp. deltoides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Hypericum densiflorum var. densiflorum", "Hypericum densiflorum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Echinochloa crusgalli var. crusgalli", "Echinochloa crus-galli", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Tiedemannia filiformis ssp.filiformis", "Oxypolis filiformis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Symphyotrichum racemosum var. racemosum", "Symphyotrichum racemosum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Rorippa palustris ssp.palustris", "Rorippa palustris ssp. palustris", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Diaperia verna var. verna", "Evax verna var. verna", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Luzula multiflora var. multiflora", "Luzula multiflora ssp. multiflora var. multiflora", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Galium orizabense ssp.laevicaule", "Galium orizabense ssp. laevicaule", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coleataenia longifolia ssp.combsii", "Panicum rigidulum var. pubescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Smallanthus uvedalia", "Smallanthus uvedalius", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Viola lanceolata var. lanceolata", "Viola lanceolata ssp. lanceolata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ludwigia linearis var. linearis", "Ludwigia linearis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium commutatum var. commutatum", "Dichanthelium commutatum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Quercus laurifolia 'waccamawensis'", "Quercus laurifolia", VegBank_Proj129$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=4911 this says synonym is Panicum curtifolium which redirects to this one
VegBank_Proj129$bestname <- gsub("Dichanthelium curtifolium", "Dichanthelium acuminatum var. fasciculatum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Saccharum strictus", "Saccharum baldwinii", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Juncus scirpoides var. scirpoides", "Juncus scirpoides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Polygonum polygamum var. polygamum", "Polygonella polygama", VegBank_Proj129$bestname)

#http://newyork.plantatlas.usf.edu/Plant.aspx?id=1450
VegBank_Proj129$bestname <- gsub("Hylodesmum pauciflorum", "Desmodium pauciflorum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Polygonum polygamum var. croomii", "Polygonella polygama", VegBank_Proj129$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3440
VegBank_Proj129$bestname <- gsub("Viola edulis", "Viola palmata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Lechea pulchella var. ramosissima", "Lechea pulchella", VegBank_Proj129$bestname)

#https://www.missouribotanicalgarden.org/PlantFinder/PlantFinderDetails.aspx?taxonid=287531&isprofile=0&n=1 
VegBank_Proj129$bestname <- gsub("Viola eriocarpa", "Viola pubescens var. pubescens", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Chasmanthium sessiliflorum var. sessiliflorum", "Chasmanthium sessiliflorum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex bromoides ssp.bromoides", "Carex bromoides ssp. bromoides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Andropogon dealbatus", "Andropogon capillipes", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Viola tripartita var. tripartita", "Viola tripartita", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Luziola fluitans var. fluitans", "Luziola fluitans", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Lupinus perennis ssp.perennis", "Lupinus perennis ssp. perennis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ludwigia grandiflora ssp.hexapetala", "Ludwigia grandiflora", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Urtica dioica ssp.dioica", "Urtica dioica ssp. dioica", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium portoricense ssp.portoricense", "Dichanthelium sabulorum var. thinium ", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Dichanthelium sabulorum var. thinium ", "Dichanthelium sabulorum var. thinium", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coleataenia anceps ssp.rhizomata", "Panicum anceps", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Lespedeza hirta var. hirta", "Lespedeza hirta ssp. hirta", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Paspalum praecox var. praecox", "Paspalum praecox", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coleataenia anceps ssp.anceps", "Panicum anceps", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coleataenia rigidula ssp.condensa", "Panicum rigidulum", VegBank_Proj129$bestname)


VegBank_Proj129$bestname <- gsub("Viola acaulescent", "Viola", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Poa trivialis ssp.trivialis", "Poa trivialis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Rubus Dewberry", "Rubus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Betula lenta var. lenta", "Betula lenta", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Circaea alpina ssp.alpina", "Circaea alpina ssp. alpina", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Pteridium latiusculum var. latiusculum", "Pteridium aquilinum var. latiusculum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Magnolia acuminata var. acuminata", "Magnolia acuminata", VegBank_Proj129$bestname)

#https://www.carolinanature.com/trees/rhpi.html
VegBank_Proj129$bestname <- gsub("Rhododendron pilosum", "Menziesia pilosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Silphium asteriscus var. trifoliatum", "Silphium asteriscus", VegBank_Proj129$bestname)

#http://newyork.plantatlas.usf.edu/Plant.aspx?id=1808
VegBank_Proj129$bestname <- gsub("Anticlea glauca", "Zigadenus elegans ssp. elegans", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/homalosorus-pycnocarpos/
VegBank_Proj129$bestname <- gsub("Homalosorus pycnocarpos", "Diplazium pycnocarpon", VegBank_Proj129$bestname)

#https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi?id=57934
VegBank_Proj129$bestname <- gsub("Sibbaldia retusa", "Sibbaldiopsis tridentata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Triphora trianthophoros var. trianthophoros", "Triphora trianthophora", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Erythronium americanum ssp.americanum", "Erythronium americanum ssp. americanum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Calystegia spithamaea ssp.stans", "Calystegia spithamaea ssp. stans", VegBank_Proj129$bestname)

#http://www.namethatplant.net/plantdetail.shtml?plant=692
VegBank_Proj129$bestname <- gsub("Pilosella caespitosa", "Hieracium caespitosum", VegBank_Proj129$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=5441
VegBank_Proj129$bestname <- gsub("Carex austrolucorum", "Carex lucorum var. austrolucorum", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/convallaria-pseudomajalis/
VegBank_Proj129$bestname <- gsub("Convallaria pseudomajalis", "Convallaria majuscula", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Osmunda claytoniana var. claytoniana", "Osmunda claytoniana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Myriopteris tomentosa", "Cheilanthes lanosa", VegBank_Proj129$bestname)

#http://newyork.plantatlas.usf.edu/plant.aspx?id=787
VegBank_Proj129$bestname <- gsub("Mononeuria groenlandica", "Minuartia groenlandica", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Phleum pratense ssp.pratense", "Phleum pratense", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Poa pratensis ssp.pratensis", "Poa pratensis ssp. pratensis", VegBank_Proj129$bestname)

#https://plants.ncwildflower.org/plant_galleries/details/huperzia-appressa
VegBank_Proj129$bestname <- gsub("Huperzia appressa", "Huperzia appalachiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Arabidopsis lyrata ssp.lyrata", "Arabis lyrata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Alnus viridis ssp.crispa", "Alnus viridis ssp. crispa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Asplenium trichomanes ssp.trichomanes", "Asplenium trichomanes ssp. trichomanes", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Deschampsia cespitosa ssp.glauca", "Deschampsia cespitosa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Oreojuncus trifidus", "Juncus trifidus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Vicia sativa ssp.nigra", "Vicia sativa ssp. nigra", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/primula-meadia/
VegBank_Proj129$bestname <- gsub("Primula meadia", "Dodecatheon meadia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Woodsia obtusa ssp.obtusa", "Woodsia obtusa ssp. obtusa", VegBank_Proj129$bestname)

#There is no Dioscorea species really close to polystachya but this species has the same common name
VegBank_Proj129$bestname <- gsub("Dioscorea polystachya", "Dioscorea oppositifolia", VegBank_Proj129$bestname)

#http://newyork.plantatlas.usf.edu/Plant.aspx?id=423
VegBank_Proj129$bestname <- gsub("Pilosella piloselloides", "Hieracium piloselloides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Paspalum dilatatum ssp.dilatatum", "Paspalum dilatatum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Packera paupercula var. appalachiana", "Packera paupercula", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex echinata ssp.echinata", "Carex echinata ssp. echinata", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Saccharum contortus", "Saccharum brevibarbe var. contortum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Holcus mollis ssp.mollis", "Holcus mollis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Symphyotrichum concinnum", "Symphyotrichum laeve var. concinnum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Elymus canadensis var. canadensis", "Elymus canadensis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Lycopodioides apodum", "Selaginella apoda", VegBank_Proj129$bestname)

#https://plants.ncwildflower.org/plant_galleries/details/cleistes-bifaria
VegBank_Proj129$bestname <- gsub("Cleistesiopsis bifaria", "Cleistes bifaria", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Myosotis laxa ssp.laxa", "Myosotis laxa", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex leptalea var. leptalea", "Carex leptalea ssp. leptalea", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex bromoides ssp.montana", "Carex bromoides ssp. montana", VegBank_Proj129$bestname)

#http://floraofalabama.org/Plant.aspx?id=4137
VegBank_Proj129$bestname <- gsub("Cyperus hortensis", "Kyllinga pumila", VegBank_Proj129$bestname)

#https://plants.ces.ncsu.edu/plants/pilosella-officinarum/
VegBank_Proj129$bestname <- gsub("Pilosella officinarum", "Hieracium pilosella", VegBank_Proj129$bestname)

#http://floraofalabama.org/Plant.aspx?id=1208
VegBank_Proj129$bestname <- gsub("Packera crawfordii", "Packera paupercula", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Antennaria parlinii ssp.fallax", "Antennaria parlinii ssp. fallax", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Sarracenia purpurea var. montana", "Sarracenia purpurea ssp. purpurea var. montana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Helianthus occidentalis ssp.occidentalis", "Helianthus occidentalis ssp. occidentalis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Campanula aparinoides var. aparinoides", "Campanula aparinoides", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Vaccinium stamineum var. stamineum", "Vaccinium stamineum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Physostegia virginiana ssp.virginiana", "Physostegia virginiana ssp. virginiana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Antennaria howellii ssp.petaloidea", "Antennaria howellii ssp. petaloidea", VegBank_Proj129$bestname)

#http://www.namethatplant.net/plantdetail.shtml?plant=4296
VegBank_Proj129$bestname <- gsub("Thalictrum hepaticum", "Thalictrum revolutum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Juncus effusus ssp.solutus", "Juncus effusus var. solutus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Packera paupercula var. paupercula", "Packera paupercula", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Fagus grandifolia var. grandifolia", "Fagus grandifolia", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Coreopsis major var. major", "Coreopsis major", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Erythronium umbilicatum ssp.monostolum", "Erythronium umbilicatum ssp. monostolum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Sphagnum spp.", "Sphagnum", VegBank_Proj129$bestname)

#these two Stellaria species listed are the only two in the genus so just reducing to genus
VegBank_Proj129$bestname <- gsub("Stellaria pubera/corei complex", "Stellaria", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Monarda fistulosa var. mollis", "Monarda fistulosa ssp. fistulosa var. mollis", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Phlox nivalis var. hentzii", "Phlox nivalis ssp. hentzii", VegBank_Proj129$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=1424
VegBank_Proj129$bestname <- gsub("Mononeuria glabra", "Minuartia glabra", VegBank_Proj129$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=3223
VegBank_Proj129$bestname <- gsub("Geum donianum", "Waldsteinia fragarioides ssp. doniana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Cyperus lupulinus var. lupulinus", "Cyperus lupulinus ssp. lupulinus", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Equisetum hyemale ssp.affine", "Equisetum hyemale var. affine", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Carex tenera var. tenera", "Carex tenera", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Solidago rugosa var. aspera", "Solidago rugosa ssp. aspera", VegBank_Proj129$bestname)

#https://gobotany.nativeplanttrust.org/species/ampelopsis/glandulosa/
VegBank_Proj129$bestname <- gsub("Ampelopsis glandulosa", "Ampelopsis brevipedunculata", VegBank_Proj129$bestname)

#http://floraofalabama.org/Plant.aspx?id=4904
VegBank_Proj129$bestname <- gsub("Dichanthelium cryptanthum", "Dichanthelium scabriusculum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Ligustrum obtusifolium var. obtusifolium", "Ligustrum obtusifolium", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Chamaelirium luteum phone", "Chamaelirium luteum", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Silene caroliniana var. caroliniana", "Silene caroliniana ssp. caroliniana", VegBank_Proj129$bestname)

VegBank_Proj129$bestname <- gsub("Myriopteris", "Cheilanthes", VegBank_Proj129$bestname)


#	Could not find solid substitutes for Stachys matthewsii and Prunus speciosa. Also the genus Periscaria. It looks like some Persicaria are considered Polygonums but maybe not all? 

```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}



#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_Proj129, KPEACH_ExoticStatus_authority, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "Unknown") %>%
    filter(bestname != "Asteraceae") %>%
    filter(bestname != "Poaceae") %>%
    filter(bestname != "Fabaceae") %>%
    filter(bestname != "Lamiaceae") %>%
     filter(bestname != "Eudicotyledonae") %>%
     filter(bestname != "Cyperaceae") %>%
     filter(bestname != "Lycopodiophyta") %>%
     filter(bestname != "Algae") %>%
    filter(bestname != "Monilophyta") %>%
   filter(bestname != "Onagraceae") %>%
     filter(bestname != "Monocotyledonae") %>%
       filter(bestname != "Carex section Acrocystis") %>%
       filter(bestname != "Carex section Ovales") %>%
       filter(bestname != "Poaceae tribe Paniceae") %>%
   filter(bestname != "Carex section Phaestoglochin") %>%
    filter(bestname != "Caryophyllaceae") %>%
    filter(bestname != "Rosaceae") %>%
    filter(bestname != "Plantae") %>%
      filter(bestname != "Orchidaceae") %>%
      filter(bestname != "Polygonaceae") %>%
      filter(bestname != "Brassicaceae") %>%
      filter(bestname != "Saxifragaceae") %>%
      filter(bestname != "Carex section Vulpinae") %>%
      filter(bestname != "Carex section Careyanae") %>%
        filter(bestname != "Carex section Griseae") %>%
        filter(bestname != "Carex section Lupulinae") %>%
        filter(bestname != "Carex section Hymenochlaenae") %>%
        filter(bestname != "Carex section Laxiflorae") %>%
    filter(bestname != "Apiaceae") %>%
   filter(bestname != "Apocynaceae") %>%
    filter(bestname != "Nabalus") %>% 
 filter(!str_detect(bestname, "\\d")) 

#only removing Nabalus here because I cannot really figure out what it should be. The internet says that it used to be considered part of the genus Prenanthes but there are also observations of Prenanthes at this site so if I renamed it Prenanthes it would just look like duplicated in the data table 

```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_Proj129", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```

Adding species codes to rows with matches

```{r}


VegBank_with_SpCode_Proj1291 <- dplyr::inner_join(VegBank_Proj129, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_Proj1291 <- VegBank_with_SpCode_Proj1291 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_Proj129 <- merge(VegBank_with_SpCode_Proj1291, no_match, all = TRUE)

VegBank_with_SpCode_Proj129 <- VegBank_with_SpCode_Proj129 %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species.

```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_Proj129$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_Proj129$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_Proj129$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_Proj129$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_Proj129 <- VegBank_with_SpCode_Proj129 %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))

```

Renaming USDA_Duration column to just 'Duration' 

```{r}

VegBank_Proj1295 <- VegBank_with_SpCode_Proj129 %>% rename('Duration' = 'USDA_Duration') %>% select(-Synonym.Symbol)


```

Exporting full data table

```{r}

#As csv
write.csv(VegBank_Proj1295,"/home/shares/neon-inv/output_files/VegBank_Proj129_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

VegBank_Proj1296 <- VegBank_Proj1295 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_Proj1296 <- VegBank_Proj1296 %>% drop_na(Lat)

#As csv
write.csv(VegBank_Proj1296,"/home/shares/neon-inv/output_files/VegBank_Proj129_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_Proj1296,"VegBank_Proj129_KPEACH_reduced.csv", row.names = FALSE)

```