---
title: "VegBank_AL_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Alabama VegBank Data (from a zip file delivered by email from Bob Peet)

Import Data

```{r, message=FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_AL <- read_excel(file.path(data_raw, "CVS-ALplots.xlsx"), sheet = "plot data", guess_max = min(4000, n_max = NULL))
VegBank_plot_species_list_AL <- read_excel(file.path(data_raw, "CVS-ALplots.xlsx"),  sheet = "plot species list", guess_max = min(4000, n_max = NULL))

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")


```

Extra cleanup

```{r}

VegBank_plot_data_AL$`Observation Start Date` <- gsub("UTC", " ", VegBank_plot_data_AL$`Observation Start Date`)

#unique(VegBank_plot_data_AL$`Observation Start Date`)

VegBank_plot_data_AL <- VegBank_plot_data_AL %>% drop_na(`Observation Start Date`)

VegBank_plot_data_AL <- VegBank_plot_data_AL %>%
   filter(str_detect(`Taxon Observation Area`, "\\d")) 

VegBank_plot_species_list_AL <- VegBank_plot_species_list_AL %>% drop_na(currentTaxonNameWithSp)

VegBank_plot_species_list_AL$cover <- as.numeric(VegBank_plot_species_list_AL$cover)

unique(VegBank_plot_species_list_AL$cover)

VegBank_plot_species_list_AL <- VegBank_plot_species_list_AL %>% drop_na(cover)


```

There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the data before I even really touch it.

3 Duplicate rows are in the data before I do anything to it.

```{r}

duplicates1 <- VegBank_plot_species_list_AL %>% group_by(authorObsCode, currentTaxonNameWithSp) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) %>%  filter(currentTaxonNameWithSp != "{Unknown}") 

```


Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}

VegBank_plot_data_AL$`Sampling Level` <- as.numeric(VegBank_plot_data_AL$`Sampling Level`)

VegBank_plot_data_AL2 <- VegBank_plot_data_AL %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)

VegBank_plot_species_list_AL2 <- VegBank_plot_species_list_AL %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_AL <- full_join(VegBank_plot_data_AL2, VegBank_plot_species_list_AL2, by = 'VegBankUniqueID')

#Check to make sure only Sampling Levels 4 and 5 are included
unique(VegBank_AL$`Sampling Level`)

```

Misc Cleanup

```{r}

#Reducing sampling date to just the year to match NEON data
VegBank_AL <- VegBank_AL %>% 
  mutate('Year' = substr(VegBank_AL$Year, 1, 4))

#Adding dataset and zone columns. The dataset names in the existing NEON database are pretty broad (ex. 'BLM') so VegBank seems ok but I could always add the region ID. VegBank_AK instead of VegBank if we want more precision here
VegBank_AL <- VegBank_AL %>% add_column(Dataset = 'VEGBANK_AL', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

```


```{r}


#Making exotic status compatible to NEON data
VegBank_AL$ExoticStatus <- gsub("native", "N", VegBank_AL$ExoticStatus)
VegBank_AL$ExoticStatus <- gsub("NATIVE", "N", VegBank_AL$ExoticStatus)
VegBank_AL$ExoticStatus <- gsub("exotic", "I", VegBank_AL$ExoticStatus)
VegBank_AL$ExoticStatus <- gsub("EXOTIC", "I", VegBank_AL$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_AL$Woodiness <- gsub("Y", "Woody", VegBank_AL$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_AL <- VegBank_AL  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))


#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_AL$bestname <- gsub("subsp.", "ssp.", VegBank_AL$bestname)
VegBank_AL$bestname <- gsub("subspecies", "ssp.", VegBank_AL$bestname)
VegBank_AL$bestname <- gsub("ssp..", "ssp.", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_AL$bestname, ignore.case = FALSE)

VegBank_AL$bestname <- gsub("\\bSP\\.\\b", " ", VegBank_AL$bestname)

#strip white space
VegBank_AL$bestname  <- trimws(VegBank_AL$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_AL <- VegBank_AL %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_AL$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_AL$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_AL$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_AL$bestname)

#Save as csv for Jeff to inspect
#write.csv(VegBank_multiple_species_AL ,"/home/shares/neon-inv/output_files/multiple_sp_AL_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows from the main data set until I can find a solution. 
```{r}

VegBank_AL$bestname  <- str_replace_all(VegBank_AL$bestname , "\\*|\\[|\\]", "")

```

Manually fixing some species names

First name = incorrect
second name = replacement/correct name

```{r}

VegBank_AL$bestname <- gsub("Hypericum virginicum", "Triadenum virginicum", VegBank_AL$bestname)

#There is no Lithospermum species even close to "decipiens" so reducing to genus
VegBank_AL$bestname <- gsub("Lithospermum decipiens", "Lithospermum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans ssp. radicans", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Muscadinia rotundifolia", "Vitis rotundifolia", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Nymphaea odorata ssp.odorata", "Nymphaea odorata ssp. odorata", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Hypericum walteri", "Triadenum walteri", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Lorinseria areolata", "Woodwardia areolata", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Xyris iridifolia", "Xyris laxifolia var. iridifolia", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Pteridium latiusculum", "Pteridium aquilinum var. latiusculum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Fagus grandifolia var. caroliniana", "Fagus grandifolia", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Hamamelis virginiana var. virginiana", "Hamamelis virginiana", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Solidago caesia var. caesia", "Solidago caesia", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Betula lenta var. lenta", "Betula lenta", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Athyrium asplenioides", "Athyrium filix-femina ssp. asplenioides", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Parathelypteris noveboracensis", "Thelypteris noveboracensis", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Osmundastrum cinnamomeum", "Osmunda cinnamomea", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Sceptridium dissectum", "Botrychium dissectum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Hylodesmum nudiflorum", "Desmodium nudiflorum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Erianthus", "Saccharum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Anemone acutiloba", "Hepatica nobilis var. acuta", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Carex section Ovales", "Carex", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Endodeca serpentaria", "Aristolochia serpentaria", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Cornus stricta", "Cornus foemina", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Persicaria virginiana", "Polygonum virginianum", VegBank_AL$bestname)	

VegBank_AL$bestname <- gsub("Sambucus canadensis", "Sambucus nigra", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Lobelia georgiana", "Lobelia amoena var. glandulifera", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Osmunda spectabilis", "Osmunda regalis var. spectabilis", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Anchistea virginica", "Woodwardia virginica", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Pteridium aquilinum var. latiusculum var. pseudocaudatum", "Pteridium aquilinum var. latiusculum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Hypericum stragulum", "Hypericum hypericoides ssp. multicaule", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Clitoria mariana var. mariana", "Clitoria mariana", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Anemone americana", "Hepatica nobilis var. obtusa", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Cnidoscolus stimulosus", "Cnidoscolus urens var. stimulosus", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Chasmanthium sessiliflorum var. sessiliflorum", "Chasmanthium sessiliflorum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Carex leptalea var. harperi", "Carex leptalea ssp. harperi", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Halesia diptera var. diptera", "Halesia diptera", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Mononeuria glabra", "Minuartia glabra", VegBank_AL$bestname)

#There is no Opuntia mesacantha and too many similar specific epithets to know which one they meant
VegBank_AL$bestname <- gsub("Opuntia mesacantha", "Opuntia", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Silphium asteriscus var. trifoliatum", "Silphium asteriscus", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Nabalus autumnalis", "Prenanthes autumnalis", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Magnolia acuminata var. acuminata", "Magnolia acuminata", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Dioscorea polystachya", "Dioscorea", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Nabalus trifoliolatus", "Prenanthes trifoliolata", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Viola pedata var. pedata", "Viola pedata", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Asclepias tuberosa var. rolfsii", "Asclepias tuberosa ssp. rolfsii", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Crataegus aprica", "Crataegus flava", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Persicaria punctata", "Polygonum punctatum var. punctatum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Andropogon gerardi", "Andropogon gerardii", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Houstonia nigricans", "Stenaria nigricans var. nigricans", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Diodella teres", "Diodia teres var. teres", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Quercus margarettae", "Quercus margaretta", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Trilisa odoratissima", "Carphephorus odoratissimus", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Galium obtusum var. filifolium", "Galium obtusum ssp. filifolium", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Isotrema tomentosum", "Aristolochia tomentosa", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Symphyotrichum concolor var. concolor", "Symphyotrichum concolor", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Hylodesmum pauciflorum", "Desmodium pauciflorum", VegBank_AL$bestname)

VegBank_AL$bestname <- gsub("Aureolaria levigata", "Aureolaria laevigata", VegBank_AL$bestname)


```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_AL, KPEACH_ExoticStatus_authority, by="bestname")

no_match <- no_match %>% drop_na(bestname)


to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "{Unknown}") %>%
 filter(!str_detect(bestname, "\\d"))

```

```{r}

no_match <- unique(no_match)

```


Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_AL", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```


Adding species codes to rows with matches

```{r}

VegBank_AL <- unique(VegBank_AL)

VegBank_with_SpCode_AL1 <- dplyr::inner_join(VegBank_AL, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_AL1 <- VegBank_with_SpCode_AL1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_AL <- merge(VegBank_with_SpCode_AL1, no_match, all = TRUE)

VegBank_with_SpCode_AL <- VegBank_with_SpCode_AL %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')


```

```{r}

#Save 
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_AL_KPEACH.csv", row.names = FALSE)

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 


```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_AL$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_AL$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_AL$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_AL$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_AL <- VegBank_with_SpCode_AL %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))



```



Renaming USDA_Duration column to just 'Duration' per Lais' request

```{r}

VegBank_AL5 <- VegBank_with_SpCode_AL %>% rename('Duration' = 'USDA_Duration') %>% select(-Synonym.Symbol)


```

Looking for source of duplicates

```{r}

#2 rows that are identical except for cover value

duplicates <- VegBank_AL5 %>% group_by(VegBankUniqueID, SpCode) %>% summarise(n_obs = n()) %>% filter(n_obs > 1)

```

```{r}

#As csv
write.csv(VegBank_AL5,"/home/shares/neon-inv/output_files/VegBank_AL_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

```{r}

VegBank_AL6 <- VegBank_AL5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_AL6 <- VegBank_AL6 %>% drop_na(Lat)


#As csv
write.csv(VegBank_AL6,"/home/shares/neon-inv/output_files/VegBank_AL_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_AL6,"VegBank_AL_KPEACH_reduced.csv", row.names = FALSE)

```