---
title: "VegBank_AK_KPeach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages 

```{r, message=FALSE, warning=FALSE}
library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)
library(here)
library(stringdist)

```

## Import Data

```{r, message = FALSE}
# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#Import existing NEON database from shared folder
#PAINLES_25Oct2020 <- read_csv(file.path(data_raw,"PAINLES_25Oct2020.csv"),   
 #                             col_types = cols(Site = col_character(),
 #                                              NEONvst_status = col_character())
 #                             )

#New file path to shared folder
VegBank_plot_data_AK <- read_excel(file.path(data_raw, "CVS-AKplots.xlsx"), sheet = "plot data")
VegBank_plot_species_list_AK <- read_excel(file.path(data_raw, "CVS-AKplots.xlsx"),  sheet = "plot species list")

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv(file.path(data_output,"KPEACH_ExoticStatus_authority.csv"))

```


Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet and renaming the columns to match up.

I have in my meeting notes that I should only use rows with a "Sampling Level" of 4 or higher. 

```{r}

VegBank_plot_data_AK$`Sampling Level` <- as.numeric(VegBank_plot_data_AK$`Sampling Level`)

unique(VegBank_plot_data_AK$`Sampling Level`)

VegBank_plot_data_AK2 <- VegBank_plot_data_AK %>%
  select("Author Observation Code", "Observation Start Date", "Real Latitude", "Real Longitude", "Public Latitude", "Public Longitude", "Taxon Observation Area", "Sampling Level", "previousObsCode") %>%
  rename( "VegBankUniqueID" = "Author Observation Code","Year" = "Observation Start Date", "Lat" = "Real Latitude", "Long" = "Real Longitude") %>%
  filter(`Sampling Level` > 3)  


VegBank_plot_species_list_AK2 <- VegBank_plot_species_list_AK %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_AK <- left_join(VegBank_plot_data_AK2, VegBank_plot_species_list_AK2, by = 'VegBankUniqueID')


#Check to make sure only Sampling Levels 4 and 5 are included
unique(VegBank_AK$`Sampling Level`)



```

Misc Cleanup

```{r}

# Reducing sampling date to just the year to match NEON data
VegBank_AK <- VegBank_AK %>% 
  mutate('Year' = substr(VegBank_AK$Year, 1, 4))

#Adding dataset and zone columns. The dataset names in the existing NEON database are pretty broad (ex. 'BLM') so VegBank seems ok, but I could always add the region ID. VegBank_AK instead of VegBank if we want more precision here
VegBank_AK <- VegBank_AK %>% 
  add_column(Dataset = 'VEGBANK', .before = "VegBankUniqueID")  %>% 
  add_column(Zone = 'L48', .before = "VegBankUniqueID")

```


Recoding "Native" and "Exotic" statuses in the VegBank data to match the NEON data

*to self- probably could concatenate within gsub to reduce this?

```{r}

#Making exotic status compatible to NEON data
VegBank_AK$ExoticStatus <- gsub("native", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("NATIVE", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("exotic", "I", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("EXOTIC", "I", VegBank_AK$ExoticStatus)

```

Recoding "Y" (which indicates, YES this species is woody) to "Woody" to match the NEON data and the GrowthForm column

```{r}

#Making the Woodiness column math the NEON 
VegBank_AK$Woodiness <- gsub("Y", "Woody", VegBank_AK$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous"". 
VegBank_AK <- VegBank_AK  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))



```


Cleanup to prevent as many 'no match' rows as possible

```{r}

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_AK$bestname <- gsub("subsp.", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("subspecies", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("ssp..", "ssp.", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_AK$bestname, ignore.case = FALSE)

#strip white space
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Making a data table with taxonomic names that include multiple specific epithets

The USDA Plant list includes SOME hybrids so we want to make sure we've cleaned up the rows that list multiple specific epithets or multiple genera to maximize the probability of finding a match with the NEON/USDA plant list

```{r}

VegBank_multiple_species_AK <- VegBank_AK %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_AK$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_AK$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_AK$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_AK$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_AK ,"/home/shares/neon-inv/output_files/multiple_sp_AK_KPEACH.csv", row.names = FALSE)


```

Removing the brackets from these rows from the main data set until I can find a solution. 
```{r}

VegBank_AK$bestname  <- str_replace_all(VegBank_AK$bestname , "\\*|\\[|\\]", "")

```


Manually fixing some species. A quick search on USDA's website (https://plants.usda.gov/) can usually tell me what they correct name should be but sometimes it requires a little more searching. For some species that required a little more sleuthing I included a commented out link to a reference website that would help anyone reading this understand why I made the choice I did. 

```{r}


VegBank_AK$bestname <- gsub("Botrypus virginianus", "Botrychium virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_AK$bestname)

#https://www.carolinanature.com/trees/trdi.html
VegBank_AK$bestname <- gsub("Thyrsanthella difformis", "Trachelospermum difforme", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Endodeca serpentaria", "Aristolochia serpentaria", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus nigra ssp.canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Muscadinia rotundifolia", "Vitis rotundifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Fagus grandifolia var. caroliniana", "Fagus grandifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Smilax hispida", "Smilax tamnoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Nabalus", "Prenanthes", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Persicaria virginiana", "Polygonum virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Populus deltoides ssp.deltoides", "Populus deltoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum", VegBank_AK$bestname)
	

#strip white space
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_AK, KPEACH_ExoticStatus_authority, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "{Unknown}") %>%
 filter(!str_detect(bestname, "\\d"))


```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_AK", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```


Adding species codes to rows with matches

```{r}


VegBank_with_SpCode_AK1 <- dplyr::inner_join(VegBank_AK, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_AK1 <- VegBank_with_SpCode_AK1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_AK <- merge(VegBank_with_SpCode_AK1, no_match, all = TRUE)

VegBank_with_SpCode_AK <- VegBank_with_SpCode_AK %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')


```

```{r}

#Save 
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_AK_KPEACH.csv", row.names = FALSE)

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 


```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_AK$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_AK$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_AK$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_AK$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_AK <- VegBank_with_SpCode_AK %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))



```


Renaming USDA_Duration column to just 'Duration' per Lais' request

```{r}

VegBank_AK5 <- VegBank_with_SpCode_AK %>% 
  rename('Duration' = 'USDA_Duration') %>% 
  select(-Synonym.Symbol)

```


```{r, eval=FALSE}

#This line is not working
write.csv(VegBank_AK5, file.path(data_output,"VegBank_AK_KPEACH_ALL_COLS.csv", row.names = FALSE))

#Manual save
write.csv(VegBank_AK5,"/home/shares/neon-inv/output_files/VegBank_AK_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

```{r, eval=FALSE}

VegBank_AK6 <- VegBank_AK5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_AK6 <- VegBank_AK6 %>% drop_na(Lat)

#As csv
write.csv(VegBank_AK6,"/home/shares/neon-inv/output_files/VegBank_AK_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_AK6,"VegBank_AK_KPEACH_reduced.csv", row.names = FALSE)

```

