---
title: "VegBank_FL_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Florida VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_FL <- read_excel(file.path(data_raw, "CVS-FLplots.xlsx"), sheet = "plot data")
VegBank_plot_species_list_FL <- read_excel(file.path(data_raw, "CVS-FLplots.xlsx"),  sheet = "plot species list")

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority.csv")

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}

VegBank_plot_data_FL$`Sampling Level` <- as.numeric(VegBank_plot_data_FL$`Sampling Level`)

VegBank_plot_data_FL2 <- VegBank_plot_data_FL %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)  

VegBank_plot_species_list_FL2 <- VegBank_plot_species_list_FL %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_FL <- left_join(VegBank_plot_data_FL2, VegBank_plot_species_list_FL2, by = 'VegBankUniqueID')


```

Misc Cleanup

```{r}

#Reducing sampling date to just the year to match NEON data
VegBank_FL <- VegBank_FL %>% 
  mutate('Year' = substr(VegBank_FL$Year, 1, 4))

#Adding dataset and zone columns. The dataset names in the existing NEON database are pretty broad (ex. 'BLM') so VegBank seems ok but I could always add the region ID. VegBank_AK instead of VegBank if we want more precision here
VegBank_FL <- VegBank_FL %>% add_column(Dataset = 'VEGBANK', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

```

```{r}

#Making exotic status compatible to NEON data
VegBank_FL$ExoticStatus <- gsub("native", "N", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("NATIVE", "N", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("exotic", "I", VegBank_FL$ExoticStatus)
VegBank_FL$ExoticStatus <- gsub("EXOTIC", "I", VegBank_FL$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_FL$Woodiness <- gsub("Y", "Woody", VegBank_FL$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_FL <- VegBank_FL  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_FL$bestname <- gsub("subsp.", "ssp.", VegBank_FL$bestname)
VegBank_FL$bestname <- gsub("subspecies", "ssp.", VegBank_FL$bestname)
VegBank_FL$bestname <- gsub("ssp..", "ssp.", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("\\bSP\\.\\b", " ", VegBank_FL$bestname)

#strip white space
VegBank_FL$bestname  <- trimws(VegBank_FL$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_FL <- VegBank_FL %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_FL$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_FL$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_FL$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_FL$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_FL ,"/home/shares/neon-inv/output_files/multiple_sp_FL_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows 
```{r}

VegBank_FL$bestname  <- str_replace_all(VegBank_FL$bestname , "\\*|\\[|\\]", "")

VegBank_FL$bestname  <- str_replace_all(VegBank_FL$bestname , "\\*|\\{|\\}", "")

```


Manually fixing some species names

```{r}

VegBank_FL$bestname <- gsub("Dichanthelium tenue", "Dichanthelium dichotomum var. tenue", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia anceps ssp.rhizomata", "Panicum anceps", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Vaccinium stamineum var. caesium", "Vaccinium stamineum", VegBank_FL$bestname)


VegBank_FL$bestname <- gsub("Coleataenia longifolia ssp.longifolia", "Coleataenia longifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Galactia volubilis var. volubilis", "Galactia volubilis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Triplasis purpurea var. purpurea", "Triplasis purpurea", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Trilisa odoratissima", "Carphephorus odoratissimus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Aristida virgata", "Aristida purpurascens var. virgata", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cyperus sp.", "Cyperus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Quercus margarettae", "Quercus margaretta", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Trilisa odoratissima", "Carphephorus odoratissimus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cnidoscolus stimulosus", "Cnidoscolus urens var. stimulosus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium angustifolium", "Dichanthelium aciculare", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Ionactis linariifolia", "Ionactis linariifolius", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Asemeia grandiflora", "Polygala grandiflora var. grandiflora", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Ruellia ciliosa", "Ruellia caroliniensis ssp. ciliosa", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Muscadinia rotundifolia", "Vitis rotundifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Morella pumila", "Morella", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Quercus elliottii", "Quercus pumila", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Pteridium latiusculum", "Pteridium aquilinum var. latiusculum", VegBank_FL$bestname)

#There is no virgata in Solidago and there are too many close options for me to make a good guess
VegBank_FL$bestname <- gsub("Solidago virgata", "Solidago", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Trilisa paniculata", "Carphephorus paniculatus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Lachnanthes caroliniana", "Lachnanthes caroliana", VegBank_FL$bestname)

#No species even close to this exist
VegBank_FL$bestname <- gsub("Andropogon dealbatus", "Andropogon", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Pluchea baccharis", "Pluchea rosea", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cladium jamaicense", "Cladium mariscus ssp. jamaicense", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Andropogon perangustatus", "Andropogon gyrans var. stenophyllus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium ensifolium", "Dichanthelium dichotomum var. ensifolium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Symphyotrichum chapmanii", "Eurybia chapmanii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Vernonia angustifolia var. mohrii", "Vernonia angustifolia ssp. mohrii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Osmundastrum cinnamomeum", "Osmunda cinnamomea", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Erianthus coarctatus", "Saccharum coarctatum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Endodeca serpentaria", "Aristolochia serpentaria", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sophronanthe hispida", "Gratiola hispida", VegBank_FL$bestname)

#Not a typo. These next TWO species moved into the same new species and variety
VegBank_FL$bestname <- gsub("Dichanthelium columbianum", "Dichanthelium sabulorum var. thinium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium portoricense", "Dichanthelium sabulorum var. thinium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Ceanothus americanus var. intermedius", "Ceanothus americanus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Orbexilum lupinellus", "Orbexilum lupinellum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Polygonum pinicola", "Carphephorus odoratissimus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Piriqueta caroliniana", "Piriqueta cistoides ssp. caroliniana", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Lespedeza hirta var. hirta", "Lespedeza hirta ssp. hirta", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans", VegBank_FL$bestname)

#https://swbiodiversity.org/seinet/taxa/index.php?taxauthid=1&taxon=216779&clid=4060
VegBank_FL$bestname <- gsub("Dichanthelium webberianum", "Dichanthelium sabulorum var. thinium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia longifolia", "Panicum rigidulum var. pubescens", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Andropogon hirsutior", "Andropogon glomeratus var. hirsutior", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Rhexia mariana var. exalbida", "Rhexia mariana var. mariana", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sarracenia minor var. minor", "Sarracenia minor", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Galactia minor", "Galactia microphylla", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Paspalum setaceum var. muhlenbergii", "Paspalum setaceum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Amphicarpum muhlenbergianum", "Amphicarpum muehlenbergianum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Solidago chapmanii", "Solidago odora var. chapmanii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cartrema americanum", "Osmanthus americanus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium chamaelonche ssp.chamaelonche", "Dichanthelium dichotomum var. ensifolium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium commutatum var. ashei", "Dichanthelium commutatum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium fusiforme", "Dichanthelium aciculare", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Tradescantia sp.", "Tradescantia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Diodella teres", "Diodia teres var. teres", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia tenera", "Panicum tenerum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Chrysopsis hyssopifolia", "Chrysopsis gossypina ssp. hyssopifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Euthamia hirtipes", "Euthamia graminifolia var. hirtipes", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Liatris secunda", "Liatris pauciflora", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Aristida tenuispica", "Aristida purpurascens var. tenuispica", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Andropogon gerardi", "Andropogon gerardii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium sp.", "Dichanthelium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Digitaria filiformis var. filiformis", "Digitaria filiformis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Anchistea virginica", "Woodwardia virginica", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Andropogon tenuispatheus", "Andropogon glomeratus var. pumilus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cornus stricta", "Cornus foemina", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sporobolus vaseyi", "Calamovilfa curtissii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium caerulescens", "Dichanthelium dichotomum var. dichotomum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichantheliumaerocarpon", "Dichanthelium sphaerocarpon", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Osmunda spectabilis", "Osmunda regalis var. spectabilis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Vaccinium stamineum var. stamineum", "Vaccinium stamineum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Aletris sp.", "Aletris", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Silphium asteriscus var. dentatum", "Silphium asteriscus var. laevicaule", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Muhlenbergia sericea", "Muhlenbergia filipes", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Erianthus giganteus", "Saccharum giganteum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Solidago pulverulenta", "Solidago puberula var. pulverulenta", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Phyla nodiflora var. nodiflora", "Phyla nodiflora", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium lucidum", "Dichanthelium dichotomum var. dichotomum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium dichotomum var. nitidum", "Dichanthelium dichotomum var. dichotomum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Mnesithea tesselata", "Mnesithea", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Thyrsanthella difformis", "Trachelospermum difforme", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Panicum rigidulum var. pubescens ssp.combsii", "Panicum rigidulum var. pubescens", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Iris sp.", "Iris", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Chasmanthium sessiliflorum var. sessiliflorum", "Chasmanthium sessiliflorum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Marshallia tenuifolia", "Marshallia graminifolia var. cynanthera", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Tiedemannia filiformis ssp.filiformis", "Oxypolis filiformis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Symphyotrichum racemosum var. racemosum", "Symphyotrichum racemosum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Crocanthemum corymbosum", "Helianthemum corymbosum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sceptridium biternatum", "Botrychium biternatum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Andropogon cabanisii", "Andropogon ternarius var. cabanisii", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Euphorbia cordifolia", "Chamaesyce cordifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Rhynchospora glomerata var. glomerata", "Rhynchospora glomerata", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Stylisma angustifolia", "Stylisma patens ssp. angustifolia", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Asimina pygmaea", "Asimina pygmea", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Lithospermum virginianum", "Onosmodium virginianum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia rigidula ssp.rigidula", "Panicum rigidulum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Hymenachne hemitomon", "Hymenachne", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Bryodesma arenicola ssp.arenicola", "Selaginella arenicola ssp. arenicola", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Hydrocotyle sp.", "Hydrocotyle", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coreopsis sp.", "Coreopsis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Polygonum nesomii", "Polygonella robusta", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Passiflora sp.", "Passiflora", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Zamia floridana var. umbrosa", "Zamia pumila ssp. pumila", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Lycopus sp.", "Lycopus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Aristida rhizomorpha", "Aristida rhizomophora", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sporobolus bakeri", "Spartina bakeri", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Coleataenia rigidula ssp.condensa", "Panicum rigidulum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichantheliumetum", "Dichanthelium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Myriophyllum sp.", "Myriophyllum", VegBank_FL$bestname)



VegBank_FL$bestname <- gsub("Carphephorus odoratissimus var. subtropicanus", "Carphephorus odoratissimus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Polygonum polygamum var. polygamum", "Polygonum polygamum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Lupinus perennis ssp.gracilis", "Lupinus perennis ssp. gracilis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium chamaelonche ssp.breve", "Dichanthelium dichotomum var. breve", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Stillingia sylvatica ssp.tenuis", "Stillingia sylvatica ssp. tenuis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Ludwigia linearis var. linearis", "Ludwigia linearis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Dichanthelium neuranthum", "Dichanthelium aciculare", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Axonopus sp.", "Axonopus", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Cirsium sp.", "Cirsium", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Polygonum polygamum var. brachystachyum", "Polygonella polygama", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Eragrostis ciliaris var. ciliaris", "Eragrostis ciliaris var. brachystachya", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Bartonia paniculata ssp.paniculata", "Bartonia paniculata ssp. paniculata", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Piriqueta cistoides ssp. caroliniana var. caroliniana", "Piriqueta cistoides ssp. caroliniana", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Tiedemannia filiformis ssp.greenmanii", "Oxypolis filiformis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Orbexilum psoralioides", "Orbexilum pedunculatum var. psoralioides", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Telmatoblechnum serrulatum", "Blechnum serrulatum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Triodanis sp.", "Triodanis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Crinum americanum var. americanum", "Crinum americanum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Crocanthemum", "Helianthemum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Sideroxylon reclinatum ssp.reclinatum", "Sideroxylon reclinatum ssp. reclinatum", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Hypochaeris chillensis", "Hypochaeris brasiliensis", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Nymphaea odorata ssp.odorata", "Nymphaea odorata ssp. odorata", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Polygonum delopyrum", "Polygonella ciliata", VegBank_FL$bestname)

VegBank_FL$bestname <- gsub("Pteridium aquilinum var. latiusculum var. pseudocaudatum", "Pteridium aquilinum var. latiusculum", VegBank_FL$bestname)

```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_FL, KPEACH_ExoticStatus_authority, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))



```

```{r}

no_match <- unique(no_match)

```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_FL", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```

Adding species codes to rows with matches

```{r}

VegBank_FL <- unique(VegBank_FL)

VegBank_with_SpCode_FL1 <- dplyr::inner_join(VegBank_FL, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_FL1 <- VegBank_with_SpCode_FL1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_FL <- merge(VegBank_with_SpCode_FL1, no_match, all = TRUE)

VegBank_with_SpCode_FL <- VegBank_with_SpCode_FL %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')


```

```{r}

#Save 
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_FL_KPEACH.csv", row.names = FALSE)

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 

```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_FL$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_FL$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_FL$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_FL$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_FL <- VegBank_with_SpCode_FL %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))



```



Renaming USDA_Duration column to just 'Duration' 

```{r}

VegBank_FL5 <- VegBank_with_SpCode_FL %>% 
  rename('Duration' = 'USDA_Duration') %>% 
  select(-Synonym.Symbol)

```

```{r}

#As csv
write.csv(VegBank_FL5,"/home/shares/neon-inv/output_files/VegBank_FL_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

```{r}

VegBank_FL6 <- VegBank_FL5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_FL6 <- VegBank_FL6 %>% drop_na(Lat)

#As csv
write.csv(VegBank_FL6,"/home/shares/neon-inv/output_files/VegBank_FL_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_FL6,"VegBank_FL_KPEACH_reduced.csv", row.names = FALSE)

```