---
title: "VegBank_AK_KPeach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)
library(here)
library(stringdist)

```

Import Data

```{r}

#Import existing NEON database from shared folder
PAINLES_25Oct2020 <- read_csv("/home/shares/neon-inv/raw_VegBank_data/PAINLES_25Oct2020.csv")

#Import Alaska VegBank Data from shared folder

VegBank_plot_data_AK <- read_excel("/home/shares/neon-inv/raw_VegBank_data/CVS-AKplots.xlsx", sheet = "plot data")

VegBank_plot_species_list_AK <- read_excel("/home/shares/neon-inv/raw_VegBank_data/CVS-AKplots.xlsx",  sheet = "plot species list")

#Import USDA Plant List 
USDA_list_w_exotic_status <- read.delim2("/home/shares/neon-inv/raw_VegBank_data/USDA_Plant_List_020821.txt",  sep = ",")

```


Making a lookup table of exotic status for species already present in the NEON database

```{r}


getExoticStatus <- PAINLES_25Oct2020$ExoticStatus

names(getExoticStatus) <- PAINLES_25Oct2020$SpCode

#So to see the exotic status of Achillea millefolium or ACMI2 you would use this 
getExoticStatus['ACMI2']

#If you want just the value, without the key, add the unname function to that value you get back:

unname(getExoticStatus['ACMI2'])

#Also just making a table...I guess I don't understand the difference between this and the lookup table. Maybe just slightly different functionality?

NEON_exotic_status <- PAINLES_25Oct2020 %>% select('SpCode', 'ExoticStatus') %>% rename('NEON_Exotic_Status' = 'ExoticStatus')

NEON_exotic_status <- NEON_exotic_status[!duplicated(NEON_exotic_status$SpCode), ]

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet and renaming the columns to match up

```{r}


VegBank_plot_data_AK2 <- VegBank_plot_data_AK %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
  filter('Sampling Level' >  4)

VegBank_plot_species_list_AK2 <- VegBank_plot_species_list_AK %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic', 'taxon_usda_growthForm', 'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_AK <- full_join(VegBank_plot_data_AK2, VegBank_plot_species_list_AK2, by = 'VegBankUniqueID')


```

Misc Cleanup

```{r}

#Reducing sampling date to just the year to match NEON data
VegBank_AK <- VegBank_AK %>% 
  mutate('Year' = substr(VegBank_AK$Year, 1, 4))

#Adding dataset and zone columns. The dataset names in the existing NEON database are pretty broad (ex. 'BLM') so VegBank seems ok but I could always add the region ID. VegBank_AK instead of VegBank if we want more precision here
VegBank_AK <- VegBank_AK %>% add_column(Dataset = 'VEGBANK', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

```



Adding sequential numbers to the UniqueID. The current VegBank_UniqueID is assigned to multiple species observed within the same plot at the same time. We want truly unique identifiers for each observation/row. So I am adding a '-' followed by sequential numbers (basically row numbers) to the VegBank_UniqueID to create a NEON 'UniqueID'.

Julien, there is probably a cleaner way to achieve this right?

```{r}

#Adding  consecutive numbers to each UniqueID such that each row/observation has its own unique ID

#Making a column of row numbers
VegBank_AK <- cbind(rownames(VegBank_AK), data.frame(VegBank_AK, row.names=NULL))

#Adding a - so each UniqueID will be preserved it will just have a dash followed by sequential numbers
numbers <- paste0('-', VegBank_AK$`rownames(VegBank_AK)`)

#Final step
VegBank_AK <- VegBank_AK %>% 
add_column('UniqueID' = paste0(VegBank_AK$VegBankUniqueID, numbers), .before = 'Dataset') %>% 
select(-'rownames(VegBank_AK)')

```

Recoding "Native" and "Exotic" statuses in the VegBank data to match the NEON data

*to self- probably could concatenate within gsub to reduce this?

```{r}

#Making exotic status compatible to NEON data
VegBank_AK$ExoticStatus <- gsub("native", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("NATIVE", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("exotic", "I", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("EXOTIC", "I", VegBank_AK$ExoticStatus)

```

Recoding "Y" (which indicates, YES this species is woody) to "Woody" to match the NEON data and the GrowthForm column

```{r}

#Making the Woodiness column math the NEON 
VegBank_AK$Woodiness <- gsub("Y", "Woody", VegBank_AK$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_AK <- VegBank_AK  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))



```


Cleanup to prevent as many 'no match' rows as possible

```{r}

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_AK$bestname <- gsub("subsp.", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("subspecies", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("ssp..", "ssp.", VegBank_AK$bestname)


#strip white space
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_AK <- VegBank_AK %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_AK$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_AK$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_AK$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_AK$bestname)

#Save as csv for Jeff to inspect
#write.csv(VegBank_multiple_species_AK ,"/home/shares/neon-inv/output_files/multiple_sp_AK_KPEACH.csv", row.names = FALSE)

```

Removing these rows from the main data set until I can find a solution. I don't know what kind magic brackets have been 
```{r}

VegBank_AK <- VegBank_AK %>%
    filter(!str_detect(bestname,  ".\\[|\\]."))

VegBank_AK$bestname  <- str_replace_all(VegBank_AK$bestname , "\\*|\\[|\\]", "")

```

Read in USDA Plants list with species codes and only keep the native status for the L48


```{r}

#Only keeping L48 Native status by detecting string that start with L

USDA_list_w_exotic_status <- USDA_list_w_exotic_status %>%
  filter(str_detect(`Native.Status`, "L[1-9]+..")) %>% # detect row with site tag
  mutate(Native.Status = str_extract(`Native.Status`, "L[1-9]+..."))

unique(USDA_list_w_exotic_status$Native.Status)

#Reducing the column to just include that status within the parenthesis ex. L48(N) = N
USDA_list_w_exotic_status$Native.Status <- substr(USDA_list_w_exotic_status$Native.Status, 5, 7)

USDA_list_w_exotic_status$Native.Status <- gsub("[()]", "", USDA_list_w_exotic_status$Native.Status)

USDA_list_w_exotic_status$Native.Status <- gsub('[[:punct:]]', "", USDA_list_w_exotic_status$Native.Status)

#Check to make sure this worked. I can see that the native status symbols that remain are "I", "N", "NI" and "W". I will have to ask the NEON team what they want to do about the 'Waif' status individuals if they are present. There are only 16 waifs and 19 NI's so its mostly I and N species. 
unique(USDA_list_w_exotic_status$Native.Status)

summary(USDA_list_w_exotic_status$Native.Status)

#Cleaning up the table by removing some extra columns we don't need and renaming others
USDA_plant_list_ALL <- USDA_list_w_exotic_status %>%
  select(-"Symbol") %>%
  rename("bestname" = "Scientific.Name", "SpCode" = 'Accepted.Symbol', "USDA_Exotic_Status" = "Native.Status", "USDA_Growth_Form" = "Growth.Habit", "USDA_Duration" = "Duration") 



```
Reducing the USDA growth form description down to one word

```{r}

USDA_plant_list_ALL$USDA_Growth_Form <- gsub('[[:punct:]]', " ", USDA_plant_list_ALL$USDA_Growth_Form)

USDA_plant_list_ALL$USDA_Growth_Form <- stringr::word(USDA_plant_list_ALL$USDA_Growth_Form, 1)

unique(USDA_plant_list_ALL$USDA_Growth_Form)

```

The USDA Plant List, when exported from the advanced search function, does not include the genus symbol for these genera. They are on the website (which is how I got the codes) but for some reason they are not exported with the list. Other genera ARE exported (Abelia = ABELI) so I know it was not a search error. Also, species within these genera are exported with the list (ex. Viola adunca = VIAD) just not the genus itself. So I don't think a chunk of the list is being cut out or not exporting properly. 

```{r}


#Idk why it does not include some genera so I am going to fix manually
USDA_plant_list_ALL <- USDA_plant_list_ALL %>% 
add_row(bestname = 'Viola', SpCode = 'VIOLA', USDA_Exotic_Status = "NI")

USDA_plant_list_ALL <- USDA_plant_list_ALL %>% 
add_row(bestname = 'Polygonum', SpCode = 'POLYG4', USDA_Exotic_Status = "NI")

  
```



Here I am trying to match the native status codes in the NEON database with the native status codes withtin the USDA plant list. In other words, are all of the species listed as "I" in the NEON database also listed as "I" in the USDA database?

```{r}


test <- full_join(USDA_plant_list_ALL, NEON_exotic_status, by = 'SpCode')

#Making sure both columns are the same class so they can be compared
test$NEON_Exotic_Status <- as.character(test$NEON_Exotic_Status)

test$USDA_Exotic_Status <- as.character(test$USDA_Exotic_Status)

#They do not match exactly
all(test$NEON_Exotic_Status == test$USDA_Exotic_Status)

#30,747 rows do not match the USDA assignation of exotic status
Dont_match1 <- test[which(test[,'NEON_Exotic_Status'] != test[,'USDA_Exotic_Status']), ]

Dont_match2 <- Dont_match1 %>% add_count(bestname, sort = TRUE)

Dont_match <- Dont_match2[!duplicated(Dont_match2$bestname), ]

#373 species in the NEON database have a different exotic status in NEON then they do in the USDA plant list
dont_match_species <- unique(Dont_match$bestname)

#9,258 species in the NEON database
total_species <- unique(PAINLES_25Oct2020$bestname)

summary(Dont_match$NEON_Exotic_Status)

#725,513 rows DO match exactly!
matches <- test[which(test[,'NEON_Exotic_Status'] == test[,'USDA_Exotic_Status']), ]

write.csv(Dont_match,"/home/shares/neon-inv/output_files/Different_ExoticStatus_NEON_USDA_KPEACH.csv", row.names = FALSE)


```

Manually fixing some species

```{r}


VegBank_AK$bestname <- gsub("Botrypus virginianus", "Botrychium virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_AK$bestname)

#https://www.carolinanature.com/trees/trdi.html
VegBank_AK$bestname <- gsub("Thyrsanthella difformis", "Trachelospermum difforme", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Endodeca serpentaria", "Aristolochia serpentaria", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus nigra ssp.canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Muscadinia rotundifolia", "Vitis rotundifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Fagus grandifolia var. caroliniana", "Fagus grandifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Smilax hispida", "Smilax tamnoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Nabalus", "Prenanthes", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Persicaria virginiana", "Polygonum virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Populus deltoides ssp.deltoides", "Populus deltoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum", VegBank_AK$bestname)
	


#strip white space
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Table of species with no match in USDA table

```{r}

#11 rows with no match
no_match <- dplyr::anti_join(VegBank_AK,USDA_plant_list_ALL, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)


no_match <- no_match %>% drop_na(bestname)

#Save as csv for Jeff to inspect
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_AK_KPEACH.csv", row.names = FALSE)

```

Adding species codes to rows with matches

```{r}

#613

VegBank_with_SpCode_AK1 <- dplyr::inner_join(VegBank_AK, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_AK1 <- VegBank_with_SpCode_AK1 %>%
  select(-taxon_usda_growthForm, -SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

VegBank_with_SpCode_AK2 <- dplyr::inner_join(no_match, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_AK2 <- VegBank_with_SpCode_AK2 %>%
  select(-taxon_usda_growthForm, -SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

VegBank_with_SpCode_AK <- merge(VegBank_with_SpCode_AK1, VegBank_with_SpCode_AK2, all = TRUE)

VegBank_with_SpCode_AK <- VegBank_with_SpCode_AK %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')


```



```{r}


#this gives me a table with all the info I need. Three columns for "Exotic Status": VegBank Exotic Status, USDA exotic status, and NEON exotic status

VegBank_AK3 <- dplyr::left_join(VegBank_with_SpCode_AK, NEON_exotic_status, by = 'SpCode')

#VegBank_AK3 <- unique(VegBank_AK3)

#I am filling the NEW_ExoticStatus column with the values of the NEON_exotic_status column (where there are values) and for any NAs in the NEON_exotic_status column fill with the values from the USDA_exotic_status column

VegBank_AK4 <- VegBank_AK3 %>%
  mutate('NEW_ExoticStatus' = ifelse(!is.na(NEON_Exotic_Status), NEON_Exotic_Status, USDA_Exotic_Status))

#Yassssssssss!! VICTORY
unique(VegBank_AK4$NEW_ExoticStatus)


```


Making a new table, pulling out odd plot size values etc

```{r}

plot_sizes <- unique(VegBank_AK4$`Taxon Observation Area`)

odd_plot_sizes <- VegBank_AK4  %>%
  filter(`Taxon.Observation.Area` == '-1')

VegBank_AK5 <- VegBank_AK4  %>%
  filter(`Taxon.Observation.Area` != '-1')

#write.csv(odd_plot_sizes,"/home/shares/neon-inv/output_files/odd_plot_sizes_AK_KPEACH.csv", row.names = FALSE)

```

Removing columns we don't need

```{r}

VegBank_AK5 <- VegBank_AK5  %>%
  select(-'VegBank_GrowthForm', -'VegBankUniqueID')


```

Making a table for NEON working group that shows which rows from the AK dataset do not have matching USDA and NEON exotic statuses

```{r}

#46 rows do not match the USDA assignation of exotic status
dont_match_usda_neon_ExoticStatus <- VegBank_AK5[which(VegBank_AK5[,'NEON_Exotic_Status'] != VegBank_AK5[,'USDA_Exotic_Status']), ]

dont_match_usda_neon_ExoticStatus$PctCov <- as.numeric(dont_match_usda_neon_ExoticStatus$PctCov)

mismatch_names <- unique(dont_match_usda_neon_ExoticStatus$bestname)

mismatch_names
#As csv
write.csv(dont_match_usda_neon_ExoticStatus,"/home/shares/neon-inv/output_files/Different_ExoticStatus_VegBank_AK_KPEACH.csv", row.names = FALSE)

```



```{r}

#As csv
write.csv(VegBank_AK5,"/home/shares/neon-inv/output_files/VegBank_AK_KPEACH.csv", row.names = FALSE)

```

