---
title: "VegBank_AK_KPeach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages 

```{r, message=FALSE, warning=FALSE}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)
library(here)
library(stringdist)

```

## Import Data

```{r, message = FALSE}

# Paths to data on server
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#Import existing NEON database from shared folder. You don't really need to do this every time. It just helped me to use it as a template when I was writing the code
#PAINLES_25Oct2020 <- read_csv(file.path(data_raw,"PAINLES_25Oct2020.csv"),   
#                              col_types = cols(Site = col_character(),
 #                                             NEONvst_status = col_character())
  #                           )

#Importing site data from NCEAS/NEON shared folder. These two sheets are the ones that have everything I need to match the PAINLES_25Oct2020 doc but there are other sheets with environmental info etc
VegBank_plot_data_AK <- read_excel(file.path(data_raw, "CVS-AKplots.xlsx"), sheet = "plot data")
VegBank_plot_species_list_AK <- read_excel(file.path(data_raw, "CVS-AKplots.xlsx"),  sheet = "plot species list")

#Import NEW Exotic Status Authority document. 
KPEACH_ExoticStatus_authority <- read_csv(file.path(data_output,"KPEACH_ExoticStatus_authority_031621.csv"))


```

Extra cleanup. Some of the dates were followed by the letters "UTC" which we don't need. 

```{r}

VegBank_plot_data_AK$`Observation Start Date` <- gsub("UTC", "", VegBank_plot_data_AK$`Observation Start Date`)

unique(VegBank_plot_data_AK$`Observation Start Date`)

VegBank_plot_data_AK <- VegBank_plot_data_AK %>% drop_na(`Observation Start Date`)

#There should be no letters in this column, just numbers
VegBank_plot_data_AK <- VegBank_plot_data_AK %>%
   filter(str_detect(`Taxon Observation Area`, "\\d")) 

VegBank_plot_species_list_AK <- VegBank_plot_species_list_AK %>% drop_na(currentTaxonNameWithSp)

#Setting the class of the cover column to numeric so that later on I can filter out anything <.10
VegBank_plot_species_list_AK$cover <- as.numeric(VegBank_plot_species_list_AK$cover)

VegBank_plot_species_list_AK <- VegBank_plot_species_list_AK %>% drop_na(cover)

#Making sure there are no unexpected values in the cover column
unique(VegBank_plot_species_list_AK$cover)

```
There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the AK data before I even really touch it.

0 Duplicate rows are in the data before I do anything to it.

```{r}

duplicates1 <- VegBank_plot_species_list_AK %>% group_by(authorObsCode, currentTaxonNameWithSp) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) %>%  filter(currentTaxonNameWithSp != "{Unknown}") 

```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet and renaming the columns to match up.

I renamed several columns to match the column names of the NEON template data. I used the 'Real Latitude" and 'Real Longitude' as the 'Lat' and 'Long' columns. The VegBank data also had columns named 'Public Latitude' and 'Public Longitude' which I also included, but did not rename. I filtered the data to only include rows with a ‘Sampling Level’ > 3. All of the rows in the final output tables should have a Sampling Level of 4 or 5. I used the author observation code ('authorObsCode') from the VegBank data as the 'UniqueID'. This number ties the site to the environmental data.

```{r}

VegBank_plot_data_AK$`Sampling Level` <- as.numeric(VegBank_plot_data_AK$`Sampling Level`)

unique(VegBank_plot_data_AK$`Sampling Level`)

VegBank_plot_data_AK2 <- VegBank_plot_data_AK %>%
  select("Author Observation Code", "Observation Start Date", "Real Latitude", "Real Longitude", "Public Latitude", "Public Longitude", "Taxon Observation Area", "Sampling Level", "previousObsCode") %>%
  rename( "VegBankUniqueID" = "Author Observation Code","Year" = "Observation Start Date", "Lat" = "Real Latitude", "Long" = "Real Longitude") %>%
  filter(`Sampling Level` > 3)  

#While there is a column for USDA code, it is blank for all sites : ()
VegBank_plot_species_list_AK2 <- VegBank_plot_species_list_AK %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_AK <- left_join(VegBank_plot_data_AK2, VegBank_plot_species_list_AK2, by = 'VegBankUniqueID')


#Check to make sure only Sampling Levels 4 and/or 5 are included
unique(VegBank_AK$`Sampling Level`)



```

I reduced the sampling date to just include the year to match the NEON data. I added ‘Dataset’ and ‘Zone’ columns to the VegBank data so that it would match the NEON data. For all sites the ‘Zone’ was ‘L48’. The ‘Dataset’ name for all sites was ‘VegBank’ followed by the geographic abbreviation assigned to the site. For example, the Dataset name for the VegBank data from Alaska is ‘VegBank_AK’.

```{r}

#Inspect the Year column to make sure it is in the format you expect
unique(VegBank_AK$Year)

# Reduce sampling date to just the year to match NEON data
VegBank_AK <- VegBank_AK %>% 
  mutate('Year' = substr(VegBank_AK$Year, 1, 4))

#Adding dataset and zone columns. The dataset names in the existing NEON database are pretty broad (ex. 'BLM') so VegBank seems ok, but I could always add the region ID. VegBank_AK instead of VegBank if we want more precision here
VegBank_AK <- VegBank_AK %>% 
  add_column(Dataset = 'VEGBANK_AK', .before = "VegBankUniqueID")  %>% 
  add_column(Zone = 'L48', .before = "VegBankUniqueID")

```


Recoding "Native" and "Exotic" statuses in the VegBank data to match the NEON data

Anything described as "Native" in the VegBank data was given an 'N" code and anything "Exotic" was given an "I" code. Like the NEON data, the VegBank 'ExoticStatus' refers to native status in the lower 48 US states. Currently, we are not using the exotic status data that the VegBank data came with but I wanted it to at least be in a compatible format.

```{r}

#Making exotic status compatible to NEON data
VegBank_AK$ExoticStatus <- gsub("native", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("NATIVE", "N", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("exotic", "I", VegBank_AK$ExoticStatus)
VegBank_AK$ExoticStatus <- gsub("EXOTIC", "I", VegBank_AK$ExoticStatus)

```

Recoding "Y" (which indicates, YES this species is woody) to "Woody" to match the NEON data and the GrowthForm column

The Woodiness VegBank data included more variety within the herbaceous category (ex. "Herb Vine", "Herb Shrub", "Herb Shrub Vine"). These were reduced to "Herbaceous" to match the NEON binomial categorization of "Woody" vs. "Herbaceous" for this column.

```{r}

#Making the Woodiness column math the NEON 
VegBank_AK$Woodiness <- gsub("Y", "Woody", VegBank_AK$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous" in the Woodiness column". 
VegBank_AK <- VegBank_AK  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))



```


Cleanup to prevent as many 'no match' rows as possible

```{r}

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_AK$bestname <- gsub("subsp.", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("subspecies", "ssp.", VegBank_AK$bestname)
VegBank_AK$bestname <- gsub("ssp..", "ssp.", VegBank_AK$bestname)

#If a row has a bestname that is just the genus followed by "sp." it won't match with the exotic status authority document.  Ex. "Viola sp." has to be re-coded to "Viola" for it to line up with the SpCode list and get a match. So I am removing "sp." and "SP." anywhere they appear with a space before and after them. The space requirement prevents gsub from recoding "ssp." to "s", which we don't want.
VegBank_AK$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_AK$bestname, ignore.case = FALSE)

VegBank_AK$bestname <- gsub("\\bSP\\.\\b", " ", VegBank_AK$bestname, ignore.case = FALSE)

#Stripping any extra white space from the names
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Making a data table with taxonomic names that include multiple specific epithets

The USDA Plant list includes SOME hybrids so we want to make sure we've cleaned up the rows that list multiple specific epithets or multiple genera to maximize the probability of finding a match with the NEON/USDA plant list. This merits a closer look. Some hybrids are in the plant list with the 'x' with no space followed by a latin name. Ex. "Abies ×shastensis". You could probably fiddle with this a bit and find more matches. 

There were many examples of either hybridization or a lack of confidence in species ID in the VegBank data. For example, some genera were followed by 2-3 specific epithets,  ex. "Fraxinus [americana + biltmoreana + smallii]". I made a new data table (for each site) that only includes rows with multiple specific epithets. Those can be found in the output file shared folder on the NCEAS server (http://aurora.nceas.ucsb.edu/home/shares/neon-inv/output_files). For example, the csv file that has all of the rows with multiple specific epithets for the Alaska dataset are in a file named "multiple_sp_AK_KPEACH.csv". The file naming convention for all of the other sites is the same. I removed the brackets around the species and kept those rows in the datatable.

```{r}

VegBank_multiple_species_AK <- VegBank_AK %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_AK$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_AK$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrids
VegBank_multiple_species_AK$bestname <- gsub(".\\+|\\+.", "x", VegBank_multiple_species_AK$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_AK ,"/home/shares/neon-inv/output_files/multiple_sp_AK_KPEACH.csv", row.names = FALSE)


```

Removing the brackets from these rows from the main data set. They don't do anything
```{r}

VegBank_AK$bestname  <- str_replace_all(VegBank_AK$bestname , "\\*|\\[|\\]", "")


VegBank_AK$bestname  <- str_replace_all(VegBank_AK$bestname , "\\*|\\{|\\}", "")


```


Manually fixing some species. A quick search on USDA's website (https://plants.usda.gov/) can usually tell me what the correct name should be but sometimes it requires a little more searching. For some species that required a little more sleuthing I included a commented out link to a reference website that would help anyone reading this understand why I made the choice I did. 

```{r}


VegBank_AK$bestname <- gsub("Botrypus virginianus", "Botrychium virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_AK$bestname)

#https://www.carolinanature.com/trees/trdi.html
VegBank_AK$bestname <- gsub("Thyrsanthella difformis", "Trachelospermum difforme", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Endodeca serpentaria", "Aristolochia serpentaria", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Sambucus nigra ssp.canadensis", "Sambucus nigra", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Muscadinia rotundifolia", "Vitis rotundifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Fagus grandifolia var. caroliniana", "Fagus grandifolia", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Smilax hispida", "Smilax tamnoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Nabalus", "Prenanthes", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Persicaria virginiana", "Polygonum virginianum", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Populus deltoides ssp.deltoides", "Populus deltoides", VegBank_AK$bestname)

VegBank_AK$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum", VegBank_AK$bestname)


#strip white space
VegBank_AK$bestname  <- trimws(VegBank_AK$bestname, which = c("right"))

```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#41 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed) or a category too broad to have a SpCode (ex. Asteraceae)
no_match <- dplyr::anti_join(VegBank_AK, KPEACH_ExoticStatus_authority, by="bestname")

#Making sure the column class is numeric
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

#There should be no rows with NAs in the bestname column
no_match <- no_match %>% drop_na(bestname)

#The to_be_matched table is just for me to quickly see which rows have possible small errors in them. For some sites there are a lot of rows with just the taxonomic family or division. USDA does not assign SpCodes to families.
to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

#For the bigger datasets I started making these temporary "to_be_matched" tables that excluced names that were common in the data but that I knew had no code.
to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "Unknown") %>%
 filter(!str_detect(bestname, "\\d"))


```

Removing any accidental duplicates that may have come with the data

```{r}

no_match <- unique(no_match)

```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_AK", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```


Adding species codes to rows with matches

```{r}

VegBank_AK <- unique(VegBank_AK)

VegBank_with_SpCode_AK1 <- dplyr::inner_join(VegBank_AK, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_AK1 <- VegBank_with_SpCode_AK1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_AK <- merge(VegBank_with_SpCode_AK1, no_match, all = TRUE)

VegBank_with_SpCode_AK <- VegBank_with_SpCode_AK %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')


```

```{r}

#Save 
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_AK_KPEACH.csv", row.names = FALSE)

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 


```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_AK$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_AK$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_AK$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_AK$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_AK <- VegBank_with_SpCode_AK %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))



```


Renaming USDA_Duration column to just 'Duration' 

```{r}

VegBank_AK5 <- VegBank_with_SpCode_AK %>% 
  rename('Duration' = 'USDA_Duration') %>% 
  select(-Synonym.Symbol)

```

Looking for source of duplicates

```{r}

#0 duplicates

duplicates <- VegBank_AK5 %>% group_by(VegBankUniqueID, SpCode) %>% summarise(n_obs = n()) %>% filter(n_obs > 1)

```
Saving the full data table with all of the columns

```{r}

#This line is not working. It would be better if I could use those nice file paths I made at the top
#write.csv(VegBank_AK5, file.path(data_output,"VegBank_AK_KPEACH_ALL_COLS.csv", row.names = FALSE))

#Manual save
write.csv(VegBank_AK5,"/home/shares/neon-inv/output_files/VegBank_AK_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Making a reduced version that removes all of the columns used to generate the "NEW_GrowthForm" and "NEW-ExoticStatus" columns.

```{r}

VegBank_AK6 <- VegBank_AK5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_AK6 <- VegBank_AK6 %>% drop_na(Lat)


```

Saving reduced file locally and on the server

```{r}

#As csv
write.csv(VegBank_AK6,"/home/shares/neon-inv/output_files/VegBank_AK_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_AK6,"VegBank_AK_KPEACH_reduced.csv", row.names = FALSE)


```

