---
title: "VegBank_NC_Coastalplains_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages and existing NEON data to use as a template

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in North Carolina Coastal Plains VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE, echo = FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
VegBank_plot_data_NC_coastal <- read_excel(file.path(data_raw, "CVS-NCCoastlplainPlots.xlsx"), sheet = "plot data", guess_max = min(4000, n_max = NULL))
VegBank_plot_species_list_NC_coastal <- read_excel(file.path(data_raw, "CVS-NCCoastlplainPlots.xlsx"),  sheet = "plot species list", guess_max = min(4000, n_max = NULL))

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")

```
Extra cleanup

```{r}

VegBank_plot_data_NC_coastal$`Observation Start Date` <- gsub("UTC", " ", VegBank_plot_data_NC_coastal$`Observation Start Date`)

unique(VegBank_plot_data_NC_coastal$`Observation Start Date`)

VegBank_plot_data_NC_coastal <- VegBank_plot_data_NC_coastal %>% drop_na(`Observation Start Date`)

VegBank_plot_data_NC_coastal <- VegBank_plot_data_NC_coastal %>%
   filter(str_detect(`Taxon Observation Area`, "\\d")) 

VegBank_plot_species_list_NC_coastal <- VegBank_plot_species_list_NC_coastal %>% drop_na(currentTaxonNameWithSp)

VegBank_plot_species_list_NC_coastal$cover <- as.numeric(VegBank_plot_species_list_NC_coastal$cover)

unique(VegBank_plot_species_list_NC_coastal$cover)

VegBank_plot_species_list_NC_coastal <- VegBank_plot_species_list_NC_coastal %>% drop_na(cover)



```

50 duplicate rows are already in the data before I do anything to it

```{r}

#50 duplicates. Identical except for cover value

duplicates1 <- VegBank_plot_species_list_NC_coastal %>% group_by(authorObsCode, currentTaxonNameWithSp) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) %>%  filter(currentTaxonNameWithSp != "{Unknown}") 


```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}

VegBank_plot_data_NC_coastal$`Sampling Level` <- as.numeric(VegBank_plot_data_NC_coastal$`Sampling Level`)

VegBank_plot_data_NC_coastal2 <- VegBank_plot_data_NC_coastal %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Public Latitude', 'Public Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'VegBankUniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
   filter(`Sampling Level` > 3)  

VegBank_plot_species_list_NC_coastal2 <- VegBank_plot_species_list_NC_coastal %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic',  'taxon_usda_code', 'taxon_woody') %>%
  rename( 'VegBankUniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_NC_coastal <- left_join(VegBank_plot_data_NC_coastal2, VegBank_plot_species_list_NC_coastal2, by = 'VegBankUniqueID')

#Check to make sure only Sampling Levels 4 and 5 are included
unique(VegBank_NC_coastal$`Sampling Level`)

#Reducing sampling date to just the year to match NEON data
VegBank_NC_coastal <- VegBank_NC_coastal %>% 
  mutate('Year' = substr(VegBank_NC_coastal$Year, 1, 4))

#Adding a site column
VegBank_NC_coastal <- VegBank_NC_coastal %>% add_column(Dataset = 'VEGBANK_NC_Coastal', .before = "VegBankUniqueID")  %>% add_column(Zone = 'L48', .before = "VegBankUniqueID")

#Making exotic status compatible to NEON data
VegBank_NC_coastal$ExoticStatus <- gsub("native", "N", VegBank_NC_coastal$ExoticStatus)
VegBank_NC_coastal$ExoticStatus <- gsub("NATIVE", "N", VegBank_NC_coastal$ExoticStatus)
VegBank_NC_coastal$ExoticStatus <- gsub("exotic", "I", VegBank_NC_coastal$ExoticStatus)
VegBank_NC_coastal$ExoticStatus <- gsub("EXOTIC", "I", VegBank_NC_coastal$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_NC_coastal$Woodiness <- gsub("Y", "Woody", VegBank_NC_coastal$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_NC_coastal <- VegBank_NC_coastal  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))


#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_NC_coastal$bestname <- gsub("subsp.", "ssp.", VegBank_NC_coastal$bestname)
VegBank_NC_coastal$bestname <- gsub("subspecies", "ssp.", VegBank_NC_coastal$bestname)
VegBank_NC_coastal$bestname <- gsub("ssp..", "ssp.", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("\\bsp\\.\\b", " ", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("\\bSP\\.\\b", " ", VegBank_NC_coastal$bestname)

#strip white space
VegBank_NC_coastal$bestname  <- trimws(VegBank_NC_coastal$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_NC_coastal <- VegBank_NC_coastal %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_NC_coastal$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_NC_coastal$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrides
VegBank_multiple_species_NC_coastal$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_NC_coastal$bestname)

#Save as csv for Jeff to inspect
#write.csv(VegBank_multiple_species_NC_coastal ,"multiple_sp_NC_coastal_KPEACH.csv", row.names = FALSE)

```

Removing the brackets from these rows from the main data set until I can find a solution. 
```{r}

VegBank_NC_coastal$bestname  <- str_replace_all(VegBank_NC_coastal$bestname , "\\*|\\[|\\]", "")

VegBank_NC_coastal$bestname  <- str_replace_all(VegBank_NC_coastal$bestname , "\\*|\\{|\\}", "")

```

Manually fixing some species

```{r}


VegBank_NC_coastal$bestname <- gsub("Viola pedata var. pedata", "Viola pedata", VegBank_NC_coastal$bestname)


VegBank_NC_coastal$bestname <- gsub("Toxicodendron radicans var. radicans", "Toxicodendron radicans", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Cartrema americanum", "Osmanthus americanus", VegBank_NC_coastal$bestname)

#http://www.namethatplant.net/plantdetail.shtml?plant=988
VegBank_NC_coastal$bestname <- gsub("Opuntia mesacantha", "Opuntia humifusa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Hamamelis virginiana var. virginiana", "Hamamelis virginiana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Arthraxon hispidus var. hispidus", "Arthraxon hispidus", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Magnolia virginiana var. virginiana", "Magnolia virginiana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Iva frutescens var. frutescens", "Iva frutescens ssp. frutescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Gonolobus suberosus var. suberosus", "Matelea gonocarpos", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Paronychia baldwinii ssp.riparia", "Paronychia baldwinii ssp. riparia", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Aureolaria levigata", "Aureolaria laevigata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Paronychia baldwinii ssp.riparia", "Paronychia baldwinii ssp. riparia", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Hypericum mutilum var. mutilum", "Hypericum mutilum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Clitoria mariana var. mariana", "Clitoria mariana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium portoricense ssp.patulum", "Dichanthelium sabulorum var. thinium", VegBank_NC_coastal$bestname)


VegBank_NC_coastal$bestname <- gsub("Onoclea sensibilis var. sensibilis", "Onoclea sensibilis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Pleopeltis michauxiana", "Pleopeltis polypodioides ssp. michauxiana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Gonolobus suberosus var. granulatus", "Matelea gonocarpos", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Arthraxon hispidus var. hispidus", "Arthraxon hispidus", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Phoradendron leucarpum ssp.leucarpum", "Phoradendron leucarpum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Chamaecyparis thyoides var. thyoides", "Chamaecyparis thyoides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Lepidium virginicum ssp.virginicum", "Lepidium virginicum var. virginicum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Arenaria lanuginosa var. lanuginosa", "Arenaria lanuginosa ssp. lanuginosa var. lanuginosa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Verbascum thapsus ssp.thapsus", "Verbascum thapsus", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Veronica peregrina var. peregrina", "Veronica peregrina ssp. peregrina", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Festuca rubra ssp.rubra", "Festuca rubra ssp. rubra", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Paronychia baldwinii ssp.baldwinii", "Paronychia baldwinii ssp. baldwinii", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Triplasis purpurea var. purpurea", "Triplasis purpurea", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Galactia volubilis var. volubilis", "Galactia volubilis", VegBank_NC_coastal$bestname)

#https://www.itis.gov/servlet/SingleRpt/SingleRpt?search_topic=TSN&search_value=893824#null
VegBank_NC_coastal$bestname <- gsub("Seutera angustifolia", "Cynanchum angustifolium", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Galium obtusum var. obtusum", "Galium obtusum ssp. obtusum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Opuntia humifusa ssp.mesacantha", "Opuntia humifusa", VegBank_NC_coastal$bestname)

#https://www.marylandbiodiversity.com/viewSpecies.php?species=4078
VegBank_NC_coastal$bestname <- gsub("Hymenachne hemitomon", "Panicum hemitomon", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Nymphaea odorata ssp.odorata", "Nymphaea odorata ssp. odorata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Tripsacum dactyloides var. dactyloides", "Tripsacum dactyloides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Hedera helix var. helix", "Hedera helix", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Coleataenia longifolia ssp.longifolia", "Panicum rigidulum var. pubescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Pontederia cordata var. cordata", "Pontederia cordata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Oenothera fruticosa var. fruticosa", "Oenothera fruticosa ssp. fruticosa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Sagittaria latifolia var. latifolia", "Sagittaria latifolia", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_NC_coastal$bestname)

#https://www.itis.gov/servlet/SingleRpt/SingleRpt;jsessionid=301BE5D420CAB47EF9DE4C2B0476FA5E?search_topic=TSN&search_value=1061799#null
VegBank_NC_coastal$bestname <- gsub("Dichanthelium neuranthum", "Dichanthelium aciculare", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Galium tinctorium var. tinctorium", "Galium tinctorium", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Gaillardia pulchella var. drummondii", "Gaillardia pulchella", VegBank_NC_coastal$bestname)

#https://plants.ces.ncsu.edu/plants/pourthiaea-villosa/
VegBank_NC_coastal$bestname <- gsub("Pourthiaea villosa", "Photinia villosa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Erianthus", "Saccharum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Saccharum giganteus", "Saccharum giganteum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ilex decidua var. decidua", "Ilex decidua", VegBank_NC_coastal$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3967
VegBank_NC_coastal$bestname <- gsub("Solidago virgata", "Solidago stricta", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Rumex crispus ssp.crispus", "Rumex crispus ssp. crispus", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Glyceria striata var. striata", "Glyceria striata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Eragrostis secundiflora ssp.oxylepis", "Eragrostis secundiflora ssp. oxylepis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Kali soda", "Salsola kali ssp. kali", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Styrax americanus var. americanus", "Styrax americanus", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium dichotomum var. ramulosum", "Dichanthelium dichotomum var. dichotomum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Morella pumila", "Morella cerifera", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Symphyotrichum concolor var. concolor", "Symphyotrichum concolor", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Phlox nivalis var. nivalis", "Phlox nivalis ssp. nivalis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Vernonia angustifolia var. angustifolia", "Vernonia angustifolia ssp. angustifolia", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Bigelowia nudata var. nudata", "Bigelowia nudata ssp. nudata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Carex intumescens var. intumescens", "Carex intumescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Bryodesma rupestre", "Selaginella rupestris", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Bryodesma tortipilum", "Selaginella tortipila", VegBank_NC_coastal$bestname)

#https://plants.ces.ncsu.edu/plants/hemionitis-lanosa/
VegBank_NC_coastal$bestname <- gsub("Myriopteris lanosa", "Cheilanthes lanosa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Bryodesma arenicola", "Selaginella arenicola", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Zephyranthes atamasco", "Zephyranthes atamasca", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ulmus americana var. americana", "Ulmus americana", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Orbexilum psoralioides", "Orbexilum pedunculatum var. psoralioides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ionactis linariifolia", "Ionactis linariifolius", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Chamaecrista nictitans var. nictitans", "Chamaecrista nictitans ssp. nictitans", VegBank_NC_coastal$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3080
VegBank_NC_coastal$bestname <- gsub("Dichanthelium webberianum", "Dichanthelium sabulorum var. thinium", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium tenue", "Dichanthelium dichotomum var. tenue", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Andropogon gerardi", "Andropogon gerardii", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Comandra umbellata var. umbellata", "Comandra umbellata ssp. umbellata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Asclepias tuberosa var. tuberosa", "Asclepias tuberosa ssp. tuberosa", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Sarracenia rubra ssp.rubra", "Sarracenia rubra ssp. rubra", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Orbexilum lupinellus", "Orbexilum lupinellum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium chamaelonche ssp.chamaelonche", "Dichanthelium dichotomum var. breve", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Bulbostylis capillaris ssp.capillaris", "Bulbostylis capillaris ssp. capillaris", VegBank_NC_coastal$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3399
VegBank_NC_coastal$bestname <- gsub("Lithospermum virginianum", "Onosmodium virginianum", VegBank_NC_coastal$bestname)

#https://swbiodiversity.org/seinet/taxa/index.php?taxon=100793
VegBank_NC_coastal$bestname <- gsub("Bryodesma acanthonota", "Selaginella acanthonota", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Sarracenia minor var. minor", "Sarracenia minor", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Rhynchospora glomerata var. glomerata", "Rhynchospora glomerata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Maianthemum racemosum ssp.racemosum", "Maianthemum racemosum ssp. racemosum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Populus deltoides ssp.deltoides", "Populus deltoides ssp. deltoides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Hypericum densiflorum var. densiflorum", "Hypericum densiflorum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Echinochloa crusgalli var. crusgalli", "Echinochloa crus-galli", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Tiedemannia filiformis ssp.filiformis", "Oxypolis filiformis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Symphyotrichum racemosum var. racemosum", "Symphyotrichum racemosum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Rorippa palustris ssp.palustris", "Rorippa palustris ssp. palustris", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Diaperia verna var. verna", "Evax verna var. verna", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Luzula multiflora var. multiflora", "Luzula multiflora ssp. multiflora var. multiflora", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Galium orizabense ssp.laevicaule", "Galium orizabense ssp. laevicaule", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Coleataenia longifolia ssp.combsii", "Panicum rigidulum var. pubescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Smallanthus uvedalia", "Smallanthus uvedalius", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Viola lanceolata var. lanceolata", "Viola lanceolata ssp. lanceolata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ludwigia linearis var. linearis", "Ludwigia linearis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium commutatum var. commutatum", "Dichanthelium commutatum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Quercus laurifolia 'waccamawensis'", "Quercus laurifolia", VegBank_NC_coastal$bestname)

#http://www.floraofalabama.org/Plant.aspx?id=4911 this says synonym is Panicum curtifolium which redirects to this one
VegBank_NC_coastal$bestname <- gsub("Dichanthelium curtifolium", "Dichanthelium acuminatum var. fasciculatum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Saccharum strictus", "Saccharum baldwinii", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Juncus scirpoides var. scirpoides", "Juncus scirpoides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Polygonum polygamum var. polygamum", "Polygonella polygama", VegBank_NC_coastal$bestname)

#http://newyork.plantatlas.usf.edu/Plant.aspx?id=1450
VegBank_NC_coastal$bestname <- gsub("Hylodesmum pauciflorum", "Desmodium pauciflorum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Polygonum polygamum var. croomii", "Polygonella polygama", VegBank_NC_coastal$bestname)

#https://florida.plantatlas.usf.edu/Plant.aspx?id=3440
VegBank_NC_coastal$bestname <- gsub("Viola edulis", "Viola palmata", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Lechea pulchella var. ramosissima", "Lechea pulchella", VegBank_NC_coastal$bestname)

#https://www.missouribotanicalgarden.org/PlantFinder/PlantFinderDetails.aspx?taxonid=287531&isprofile=0&n=1 
VegBank_NC_coastal$bestname <- gsub("Viola eriocarpa", "Viola pubescens var. pubescens", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Chasmanthium sessiliflorum var. sessiliflorum", "Chasmanthium sessiliflorum", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Carex bromoides ssp.bromoides", "Carex bromoides ssp. bromoides", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Andropogon dealbatus", "Andropogon capillipes", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Viola tripartita var. tripartita", "Viola tripartita", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Luziola fluitans var. fluitans", "Luziola fluitans", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Lupinus perennis ssp.perennis", "Lupinus perennis ssp. perennis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Ludwigia grandiflora ssp.hexapetala", "Ludwigia grandiflora", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Urtica dioica ssp.dioica", "Urtica dioica ssp. dioica", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium portoricense ssp.portoricense", "Dichanthelium sabulorum var. thinium ", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Dichanthelium sabulorum var. thinium ", "Dichanthelium sabulorum var. thinium", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Coleataenia anceps ssp.rhizomata", "Panicum anceps", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Lespedeza hirta var. hirta", "Lespedeza hirta ssp. hirta", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Paspalum praecox var. praecox", "Paspalum praecox", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Coleataenia anceps ssp.anceps", "Panicum anceps", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Coleataenia rigidula ssp.condensa", "Panicum rigidulum", VegBank_NC_coastal$bestname)


VegBank_NC_coastal$bestname <- gsub("Viola acaulescent", "Viola", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Poa trivialis ssp.trivialis", "Poa trivialis", VegBank_NC_coastal$bestname)

VegBank_NC_coastal$bestname <- gsub("Rubus Dewberry", "Rubus", VegBank_NC_coastal$bestname)

```

Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(VegBank_NC_coastal, KPEACH_ExoticStatus_authority, by="bestname")

#Filtering out any row with <.10 in the cover column
no_match$PctCov <- as.numeric(no_match$PctCov) 

#Removing all rows with less than 10% cover
no_match <- no_match %>% filter(PctCov > .10)

no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "Unknown") %>%
    filter(bestname != "Asteraceae") %>%
    filter(bestname != "Poaceae") %>%
    filter(bestname != "Fabaceae") %>%
    filter(bestname != "Lamiaceae") %>%
     filter(bestname != "Eudicotyledonae") %>%
     filter(bestname != "Cyperaceae") %>%
     filter(bestname != "Lycopodiophyta") %>%
     filter(bestname != "Algae") %>%
    filter(bestname != "Monilophyta") %>%
   filter(bestname != "Onagraceae") %>%
     filter(bestname != "Monocotyledonae") %>%
       filter(bestname != "Carex section Acrocystis") %>%
       filter(bestname != "Carex section Ovales") %>%
       filter(bestname != "Poaceae tribe Paniceae") %>%
   filter(bestname != "Carex section Phaestoglochin") %>%
  filter(bestname != "Caryophyllaceae") %>%
    filter(bestname != "Rosaceae") %>%
    filter(bestname != "Plantae") %>%
      filter(bestname != "Orchidaceae") %>%
      filter(bestname != "Polygonaceae") %>%
      filter(bestname != "Brassicaceae") %>%
      filter(bestname != "Saxifragaceae") %>%
      filter(bestname != "Carex section Vulpinae") %>%
      filter(bestname != "Carex section Careyanae") %>%
        filter(bestname != "Carex section Griseae") %>%
        filter(bestname != "Carex section Lupulinae") %>%
        filter(bestname != "Carex section Hymenochlaenae") %>%
        filter(bestname != "Carex section Laxiflorae") %>%
 filter(!str_detect(bestname, "\\d")) 


```

```{r}

no_match <- unique(no_match)

```


Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_NC_Coastal", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Taxon Observation Area' = 'Taxon.Observation.Area', 'Sampling Level' = 'Sampling.Level', 'Public Latitude' = 'Public.Latitude', 'Public Longitude' = 'Public.Longitude')

```

Adding species codes to rows with matches

```{r}

VegBank_NC_coastal <- unique(VegBank_NC_coastal)

VegBank_with_SpCode_NC_coastal1 <- dplyr::inner_join(VegBank_NC_coastal, KPEACH_ExoticStatus_authority, by="bestname")

VegBank_with_SpCode_NC_coastal1 <- VegBank_with_SpCode_NC_coastal1 %>%
  select(-SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

VegBank_with_SpCode_NC_coastal <- merge(VegBank_with_SpCode_NC_coastal1, no_match, all = TRUE)

VegBank_with_SpCode_NC_coastal <- VegBank_with_SpCode_NC_coastal %>% rename('VegBank_ExoticStatus' = 'ExoticStatus', 'VegBank_GrowthForm' = 'GrowthForm')

```

I also kept all 3 GrowthForm columns. The 'VegBank_GrowthForm' is the growth form column that came with the data from Bob Peet/Jeff Corbin. The 'NEON_GrowthForm' column came from the taxonomy_temp10_revised.csv file. The 'USDA_GrowthForm' column came from the USDA plant list. I assume that the NEON group wants me to follow the same procedure as the exotic status column and use the NEON_GrowthForm whenever possible and then the USDA_GrowthForm for any new species. However, in a previous email Eve told me to use the USDA growth form and reduce it to only one word (ex. 'Forb/herb' would become 'Forb'). There are many rows in the NEON_GrowthForm column that do not follow this rule (they are more than one word). The READ ME file says to use the GrowthForm from the taxonomy_temp10_revised.csv file but also has a bullet point saying "Rule for growth form was to take the first thing listed (e.g., ‘Subshrub, Shrub’ became ‘Subshrub’)". So I made a new column called NEW_GrowthForm that uses the NEON_GrowthForm when available but the USDA_GrowthForm for new species. 

```{r}

#Removing punctuation so I can extract only the first word from this column
VegBank_with_SpCode_NC_coastal$NEON_GrowthForm <- gsub('[[:punct:]]', " ", VegBank_with_SpCode_NC_coastal$NEON_GrowthForm)

#Extracting first word
VegBank_with_SpCode_NC_coastal$NEON_GrowthForm <- stringr::word(VegBank_with_SpCode_NC_coastal$NEON_GrowthForm, 1)

#Making new column
VegBank_with_SpCode_NC_coastal <- VegBank_with_SpCode_NC_coastal %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))

```

Renaming USDA_Duration column to just 'Duration' 

```{r}

VegBank_NC_coastal5 <- VegBank_with_SpCode_NC_coastal %>% rename('Duration' = 'USDA_Duration') %>% select(-Synonym.Symbol)



```

Looking for source of duplicates

```{r}

#42 duplicates. Identical except for cover value

duplicates <- VegBank_NC_coastal5 %>% group_by(VegBankUniqueID, SpCode) %>% summarise(n_obs = n()) %>% filter(n_obs > 1)

```



Exporting full data table

```{r}

#As csv
write.csv(VegBank_NC_coastal5,"/home/shares/neon-inv/output_files/VegBank_NC_coastal_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

VegBank_NC_coastal6 <- VegBank_NC_coastal5 %>%
  select(-'VegBank_GrowthForm', -'VegBank_ExoticStatus', -'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
VegBank_NC_coastal6 <- VegBank_NC_coastal6 %>% drop_na(Lat)

#As csv
write.csv(VegBank_NC_coastal6,"/home/shares/neon-inv/output_files/VegBank_NC_coastal_KPEACH_reduced.csv", row.names = FALSE)

write.csv(VegBank_NC_coastal6,"VegBank_NC_coastal_KPEACH_reduced.csv", row.names = FALSE)

```