---
title: "VegBank_SCPlots_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages and existing NEON data to use as a template

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)

```

Reading in Texas VegBank Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}


VegBank_plot_data_TX <- read_excel("/home/shares/neon-inv/raw_VegBank_data/CVS-TXplots.xlsx", sheet = "plot data")
VegBank_plot_species_list_TX  <- read_excel("/home/shares/neon-inv/raw_VegBank_data/CVS-TXplots.xlsx",  sheet = "plot species list")


```

Selecting the columns I will need from the VegBank data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}


VegBank_plot_data_TX2 <- VegBank_plot_data_TX  %>%
  select('Author Observation Code', 'Observation Start Date', 'Real Latitude', 'Real Longitude', 'Taxon Observation Area', 'Sampling Level', 'previousObsCode') %>%
  rename( 'UniqueID' = 'Author Observation Code','Year' = 'Observation Start Date', 'Lat' = 'Real Latitude', 'Long' = 'Real Longitude' ) %>%
  filter('Sampling Level' >  4)

VegBank_plot_species_list_TX2 <- VegBank_plot_species_list_TX %>%
  select('authorObsCode', 'currentTaxonNameWithSp', 'cover', 'taxon_growthForm_short', 'taxon_woody', 'taxon_exotic', 'taxon_usda_growthForm', 'taxon_usda_code', 'taxon_woody') %>%
  rename( 'UniqueID' = 'authorObsCode','bestname' = 'currentTaxonNameWithSp', 'PctCov' = 'cover', 'SpCode' = 'taxon_usda_code', 'ExoticStatus' = 'taxon_exotic', 'GrowthForm' = 'taxon_growthForm_short', 'Woodiness' = 'taxon_woody')

VegBank_TX <- full_join(VegBank_plot_data_TX2, VegBank_plot_species_list_TX2, by = 'UniqueID')

#Reducing sampling date to just the year to match NEON data
VegBank_TX <- VegBank_TX %>% 
  mutate('Year' = substr(VegBank_TX$Year, 1, 4))

#Adding a site column
VegBank_TX <- VegBank_TX %>% add_column(Dataset = 'VEGBANK', .before = "UniqueID")  %>% add_column(Zone = 'L48', .before = "UniqueID")

#Making exotic status compatible to NEON data
VegBank_TX$ExoticStatus <- gsub("native", "N", VegBank_TX$ExoticStatus)
VegBank_TX$ExoticStatus <- gsub("NATIVE", "N", VegBank_TX$ExoticStatus)
VegBank_TX$ExoticStatus <- gsub("exotic", "NI", VegBank_TX$ExoticStatus)
VegBank_TX$ExoticStatus <- gsub("EXOTIC", "NI", VegBank_TX$ExoticStatus)

#Making the Woodiness column math the NEON 
VegBank_TX$Woodiness <- gsub("Y", "Woody", VegBank_TX$Woodiness)

# Give any row with a GrowthForm that includes "Herb" (ex. "Herb", "Herb Vine", "Herb Shrub") a Woodiness classification of ""Herbaceous". 
VegBank_TX <- VegBank_TX  %>% 
  mutate(Woodiness = ifelse(GrowthForm %in% c("Herb", "Herb Vine", "Herb Shrub", "Herb Shrub Vine"), 'Herbaceous', Woodiness))

#This just replaces all NA's, any row that was not explicitly labeled as 'Woody" as Herbaceous which could be wrong so I did not run it
#VegBank$Woodiness<-  replace_na(VegBank$Woodiness, "Herbaceous")

#Making sure all subspecies are noted correctly. Fixing some common errors that would prevent the species name from matching with the USDA plant list

VegBank_TX$bestname <- gsub("subsp.", "ssp.", VegBank_TX$bestname)
VegBank_TX$bestname <- gsub("subspecies", "ssp.", VegBank_TX$bestname)
VegBank_TX$bestname <- gsub("ssp..", "ssp.", VegBank_TX$bestname)

#strip white space
VegBank_TX$bestname  <- trimws(VegBank_TX$bestname, which = c("right"))

```

Making a datatable with taxonomic names that include multiple specific epithets

```{r}

VegBank_multiple_species_TX <- VegBank_TX %>%
    filter(str_detect(bestname,  ".\\[|\\]."))

#Removing brackets around specific epithets because the USDA Plants list does not use them..may be preventing matches
VegBank_multiple_species_TX$bestname <- gsub(".\\[|\\]", " ", VegBank_multiple_species_TX$bestname)

#Replacing  + with x because that is how USDA Plant List denotes hybrides
VegBank_multiple_species_TX$bestname <- gsub(".\\+|\\+.", " x ", VegBank_multiple_species_TX$bestname)

#Save as csv for Jeff to inspect
write.csv(VegBank_multiple_species_TX ,"/home/shares/neon-inv/output_files/multiple_sp_TX_KPEACH.csv", row.names = FALSE)

```

Removing these rows from the main data set until I can find a solution
```{r}

VegBank_TX <- VegBank_TX %>%
    filter(!str_detect(bestname,  ".\\[|\\]."))

```

Read in USDA Plants list with species codes

```{r}


USDA_plant_list_ALL <- read.delim2("/home/shares/neon-inv/raw_VegBank_data/USDA_all_with_syn.txt",  sep = ",")

USDA_plant_list_ALL <- USDA_plant_list_ALL %>%
  rename("bestname" = "Scientific.Name", "SpCode" = 'Accepted.Symbol') 

```


Table of species with no match in USDA table

```{r}

#64 rows with no match
no_match <- dplyr::anti_join(VegBank_TX,USDA_plant_list_ALL, by="bestname")

#Reducing all rows of the no match file to just the genus so that USDA can find matches
no_match$bestname <- stringr::word(no_match$bestname, 1)

#Save as csv for Jeff to inspect
write.csv(no_match,"/home/shares/neon-inv/output_files/no_match_TX_KPEACH.csv", row.names = FALSE)

```

Adding species codes to rows with matches

```{r}

#1,846
#185,454
VegBank_with_SpCode_TX1 <- dplyr::inner_join(VegBank_TX, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_TX2 <- dplyr::inner_join(no_match, USDA_plant_list_ALL, by="bestname")

VegBank_with_SpCode_TX <- merge(VegBank_with_SpCode_TX1, VegBank_with_SpCode_TX2, all = TRUE)


VegBank_with_SpCode_TX <- VegBank_with_SpCode_TX %>%
  select(-taxon_usda_growthForm, -SpCode.x) %>%
  rename("SpCode" = "SpCode.y")

#Now lets see how many no match rows are left
#0! Yay!
no_match2 <- dplyr::anti_join(VegBank_with_SpCode_TX,USDA_plant_list_ALL, by="bestname")

```

Making a new table, pulling out odd plot size values etc

```{r}

plot_sizes <- unique(VegBank_with_SpCode_TX$`Taxon Observation Area`)

odd_plot_sizes <- VegBank_with_SpCode_TX %>%
  filter(`Taxon Observation Area` == '-1')

write.csv(odd_plot_sizes,"/home/shares/neon-inv/output_files/odd_plot_sizes_TX_KPEACH.csv", row.names = FALSE)

```

```{r}

#As csv
write.csv(VegBank_with_SpCode_TX,"/home/shares/neon-inv/output_files/VegBank_TX_KPEACH.csv", row.names = FALSE)

```