---
title: "West_Virginia_KPEACH"
output: html_document
notes: "Eve Beaury started editing 4/17/2021 to incorporate more edits (particularly to taxnonomy, which was fixed outside of R in excel and then joined to the most recent datasheet Kristen had exopoted for WV. IMPORTANT: This is saved to the nceas servor but not synced to Kristen's git (I don't have permission)"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)
library(lubridate)

```

Reading in West Virginia Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
#Yikes so many errors
#Plot_species_list_WV <- read_excel(file.path(data_raw, "West_Virginia_cover_data.xlsx"),  sheet = "Plots-Species", guess_max = min(4000, n_max = NULL))

#Saved the sheets I needed as csv files
Plot_species_list_WV <- read_csv(file.path(data_raw, "WV_species_data.csv"))

Plot_data_WV <- read_csv(file.path(data_raw, "WV_plot_data.csv"))

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")

```


Removing rows with NAs in critical columns

```{r}


Plot_data_WV <- Plot_data_WV %>% drop_na(`Survey Date`)

Plot_species_list_WV <- Plot_species_list_WV %>% drop_na(Stratum)



```

There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the data before I even really touch it.

More than 19,000 Duplicate row is in the data before I do anything to it( mostly due to the different stratum).

```{r}

duplicates1 <- Plot_species_list_WV %>% group_by(`Plot Code`, `Scientific Name`) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) 

```

Selecting the columns I will need from the data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}



Plot_data_WV2 <- Plot_data_WV %>%
  select('Plot Code', 'Former Plot Code', 'Quad Code', 'Survey Date', 'Corrected Lat', 'Corrected Long', 'X Dimension', 'Y Dimension') %>%
  rename( 'UniqueID' = 'Plot Code','Year' = 'Survey Date', 'Lat' = 'Corrected Lat', 'Long' = 'Corrected Long' ) 

Plot_species_list_WV2 <- Plot_species_list_WV %>%
  select('Plot Code', 'Scientific Name', 'Stratum', 'Real Cover', 'Uncertain ID?', 'Plant Symbol') %>%
  rename( 'UniqueID' = 'Plot Code','bestname' = 'Scientific Name', 'PctCov' = 'Real Cover', 'SpCode' = 'Plant Symbol') %>% filter('Uncertain ID?' != TRUE) %>% select(-'Uncertain ID?')

WestVirginia <- left_join(Plot_data_WV2 , Plot_species_list_WV2, by = 'UniqueID')


```

```{r}

WestVirginia <- WestVirginia %>% drop_na(Year)

unique(WestVirginia$Year)

#Fixing date format
WestVirginia$Year <- lubridate::mdy(WestVirginia$Year)

#Reducing sampling date to just the year to match NEON data
WestVirginia <- WestVirginia %>% 
  mutate('Year' = substr(WestVirginia$Year, 1, 4))

unique(WestVirginia$Year)

#Adding a site column
WestVirginia <- WestVirginia %>% add_column(Dataset = 'NCVS_WV', .before = "UniqueID")  %>% add_column(Zone = 'L48', .before = "UniqueID")

WestVirginia$bestname <- gsub("\\bsp\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSP\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bL\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname   <- str_replace_all(WestVirginia$bestname , "\\*|\\[|\\]", "")

WestVirginia$bestname   <- str_replace_all(WestVirginia$bestname , "\\*|\\{|\\}", "")

```

```{r}

unique(WestVirginia$Stratum)

WestVirginia$Stratum <- as.character(WestVirginia$Stratum)

WestVirginia <- WestVirginia %>% filter(Stratum != "T")

WestVirginia <- WestVirginia %>% filter(Stratum != "N")


```

Drop rows with 0% cover measurements and no info on the observation (basically blank cells)
```{r}
WestVirginia = WestVirginia %>% filter(!bestname=="")

```


Manually fixing names
```{r}

# remove weird charaters
WestVirginia$bestname <- gsub(' - ', "", WestVirginia$bestname)
WestVirginia$bestname <- gsub(' \\(', "", WestVirginia$bestname)
WestVirginia$bestname <- gsub('\\)', "", WestVirginia$bestname)

WestVirginia$bestname <- gsub("Carya alba (L.) Nutt. ex Ell.", "Carya tomentosa", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("CAAL27", "CATO6", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Polygonum caespitosum Blume var. longisetum (de Bruyn) A.N. Steward", "Polygonum cespitosum var. longisetum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("POCAL2", "POCEL", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Eupatorium purpureum L. var. purpureum", "Eutrochium purpureum var. purpureum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("EUPUP", "EUPUP4", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Dichanthelium dichotomum (L.) Gould ssp. dichotomum", "Dichanthelium dichotomum var. dichotomum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("DIDIDWV", "DIDID", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Acalypha gracilens Gray var. gracilens", "Acalypha gracilens", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("ACGRGWV", "ACGR2", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("\\bsnag\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSnag\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSnags\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bsnags\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSNAGS\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSNAG\\b", "", WestVirginia$bestname, ignore.case = FALSE)

#Filtering out anything listed as a snag (dead tree) 
WestVirginia <- WestVirginia %>% 
  filter(!bestname %in% c("snag", "Snag", "SNAG", "SNAGS", "snags", "Snags", "snag-", "Snags - ", "snag - "))



```


Join in names fixed in excel using USDA plants (eg, spelling errors, authors included in name, weird characters, etc.).

```{r}
names <- read_csv(file.path(data_output, "no_match_WV_EBEAURY.csv"))
glimpse(names)

# drop unfixed and pull unique
names_sub = names %>% filter(fixed=="Y") %>% select(bestname, newname, newcode) %>% distinct()

# join to datasheet
glimpse(WestVirginia)
WestVirginia_join = left_join(WestVirginia, names_sub, by="bestname") # added some rows?

```

```{r}
# Merge correct codes to one column
WestVirginia_join = WestVirginia_join %>% 
  mutate(finalcode = ifelse(is.na(WestVirginia_join$newcode), SpCode, newcode),
         finalname = ifelse(is.na(newname), bestname, newname))

# Clean up & drop repeated columns
glimpse(WestVirginia_join)
WestVirginia_clean = WestVirginia_join %>% select(-c(bestname, SpCode, newname, newcode)) %>%
  rename(bestname = finalname, SpCode = finalcode)
glimpse(WestVirginia_clean)

# Fix weird codes
WestVirginia_clean %>% filter(SpCode=="Oxalis L.")
WestVirginia_clean = WestVirginia_clean %>% mutate(SpCode = ifelse(SpCode=="Oxalis L.", "OXALI", SpCode))
WestVirginia_clean = WestVirginia_clean %>% mutate(SpCode = ifelse(SpCode=="7-Feb", "FEBR7", SpCode))
WestVirginia_clean = WestVirginia_clean %>% filter(!SpCode=="snag")


```


Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(WestVirginia_clean, KPEACH_ExoticStatus_authority, by="SpCode")

unique(no_match$SpCode) # bunch of codes missing from Kristen's exotic status file?
# wonder if these are species exclusive to this park?


```

Pull Ian's taxonomy to fill in holes
```{r}

tax <- read_csv(file.path(data_raw, "taxonomy_temp10_revised.csv"))
glimpse(tax)

# pull codes we need info for
fix = data.frame(SpCode = unique(no_match$SpCode))
fix # double check dropped the NA and snag
missing = fix %>% filter(!(SpCode %in% tax$Accepted.Symbol | SpCode %in% tax$Synonym.Symbol)) # a lot aren't in Ian's thing

# 45 codes that do not match to Ian's taxonomy and aren't in Kristen's key
# check one example of a synonym code
glimpse(tax)
tax %>% filter(Accepted.Symbol=="QUMO4") # ah! so some are in there, just synonym symbol isn't filled in

```

Update some codes using USDA plants
```{r}
codes <- read_csv("/home/shares/neon-inv/output_files/USDA_Plants_ScientificNames.csv")
glimpse(codes)
glimpse(missing)
missing = left_join(missing, codes %>% rename(SpCode = Synonym.Symbol))
# join back to WV
WestVirginia_clean = WestVirginia_clean %>% left_join(missing)
glimpse(WestVirginia_clean)
# merge
WestVirginia_clean = WestVirginia_clean %>% 
  mutate(finalcode = ifelse(is.na(WestVirginia_clean$Accepted.Symbol), SpCode, Accepted.Symbol),
         finalname = ifelse(is.na(Scientific.Name), bestname, Scientific.Name))
glimpse(WestVirginia_clean)
WestVirginia_clean = WestVirginia_clean %>% mutate(finalcode = ifelse(finalcode=="7-Feb", "FEBR7", finalcode))
WestVirginia_clean = WestVirginia_clean %>% select(-c(SpCode, bestname, Accepted.Symbol, Scientific.Name)) %>%
  rename(bestname = finalname, SpCode = finalcode)

```

Look again at missing
```{r}
no_match <- dplyr::anti_join(WestVirginia_clean, KPEACH_ExoticStatus_authority, by="SpCode")
fix = data.frame(SpCode = unique(no_match$SpCode), bestname = unique(no_match$bestname))
fix # double check dropped the NA and snag
missing = fix %>% filter(!(SpCode %in% tax$Accepted.Symbol | SpCode %in% tax$Synonym.Symbol)) # only 16!


```

Fix a few by hand
```{r}
WestVirginia_clean = WestVirginia_clean %>%
  mutate(SpCode = replace(SpCode,SpCode=="EUPUP44","EUPUP4"))
WestVirginia_clean = WestVirginia_clean %>%
  mutate(SpCode = replace(SpCode,SpCode=="FUPO","FUPO70"))
WestVirginia %>% filter(SpCode=="CHARAWV") # algae
WestVirginia %>% filter(SpCode=="MORCHWV") # fungus
# rest are ssp. that should probably be varieties, but choosing not to mess with


```

Make sure accepted symbol is in SpCode (not synonyms)
```{r}
# get unique list of synonyms
syn_unique = tax %>% select(Accepted.Symbol, Synonym.Symbol, bestname) %>%
                                   rename(SpCode = Synonym.Symbol, newname = bestname)
syn_unique = syn_unique[order(syn_unique[,2]),]
syn_unique = syn_unique[!duplicated(syn_unique$SpCode),]
# join
WestVirginia_clean = WestVirginia_clean %>% left_join(syn_unique) %>%
  mutate(SpCode = ifelse(is.na(Accepted.Symbol), SpCode, Accepted.Symbol),
         bestname = ifelse(is.na(newname), bestname, newname)) %>% select(-c(newname, Accepted.Symbol))

```


Continue (remove?)
```{r}
no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "Unknown") %>%
  filter(bestname != "snag") %>%
   filter(bestname != "snags") %>%
   filter(bestname != "SNAG") %>%
   filter(bestname != "Snag") %>%
 filter(!str_detect(bestname, "\\d"))

to_be_matched <- unique(to_be_matched)
```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)
missing = missing %>% mutate(newcode = paste0("NOMATCH_WV", 1:nrow(missing)))

#Join to WV
WestVirginia_clean2 = WestVirginia_clean %>% left_join(missing) %>%
  mutate(finalcode = ifelse(is.na(newcode), SpCode, newcode)) %>%
  select(-c(SpCode, newcode)) %>% rename(SpCode = finalcode)
glimpse(WestVirginia_clean2)

#Final step
#no_match <- no_match %>% 
#mutate('SpCode' = paste0("NOMATCH_WV", numbers)) %>% 
#select(-'rownames(no_match)') %>% 
#rename('Former Plot Code' = 'Former.Plot.Code', 'Quad Code' = 'Quad.Code', 'X Dimension' = 'X.Dimension', 'Y #Dimension' = 'Y.Dimension')

```

Joining growth form and exotic status 
```{r}

# want exotic status, growth form, duration, exotic status origin
tax_unique = tax %>% select(Accepted.Symbol, Duration, GrowthForm, inv_L48) %>% rename(SpCode = Accepted.Symbol, ExoticStatus = inv_L48) %>% distinct()
# ID and remove empty rows
tax_unique %>% group_by(SpCode) %>% summarise(dups = n()) %>% filter(dups > 1)
tax_unique %>% filter(SpCode=="ACAN11") # second row is filled in...
tax_unique = tax_unique[order(tax_unique[,1], tax_unique[,2]),]
tax_unique %>% filter(SpCode=="ACAN11") # worked
tax_unique = tax_unique[!duplicated(tax_unique$SpCode),]

WestVirginia_comp = left_join(WestVirginia_clean2, tax_unique) # check rows match WV clean


```
## EVE START HERE - need to make sure I'm pulling the right column's from Kristen's key

Check for missing info
```{r}
missing_new = WestVirginia_comp %>% filter(!(SpCode %in% tax_unique$SpCode)) %>% select(SpCode, bestname) %>%
  distinct()
unique(missing_new$SpCode)
# for these - need to pull from Kristen's key
glimpse(KPEACH_ExoticStatus_authority) # W = not introduced to U.S... so hopefully we won't have any of these

# remove duplicate bestnames in Kristen's key
KPEACH_unique = KPEACH_ExoticStatus_authority %>% select(SpCode, bestname, NEW_ExoticStatus, ExoticStatus_Origin, USDA_Duration, USDA_Growth_Form) %>% distinct()
KPEACH_unique = KPEACH_unique[!duplicated(KPEACH_unique$SpCode),]
#join
missing_new = missing_new %>% left_join(KPEACH_unique, by="SpCode")
# want to retain bestname.y, 
# join back to WV
WestVirginia_comp2 = left_join(WestVirginia_comp, missing_new %>% select(-bestname.x) %>% rename(bestname=bestname.y), by="SpCode") %>%
  mutate(bestname = ifelse(is.na(bestname.y), bestname.x, bestname.y),
         Duration = ifelse(is.na(USDA_Duration), Duration, USDA_Duration),
         ExoticStatus = ifelse(is.na(NEW_ExoticStatus), ExoticStatus, NEW_ExoticStatus),
         GrowthForm = ifelse(is.na(USDA_Growth_Form), GrowthForm, USDA_Growth_Form)) %>%
  rename(NCVS_original_scientific_name = bestname.x) %>% select(-bestname.y, USDA_Duration)
glimpse(WestVirginia_comp2)
sum(is.na(WestVirginia_comp2$ExoticStatus))

```


Making a possible Taxon Observation Area column
```{r}

WestVirginia_comp2$`X Dimension` <- as.numeric(WestVirginia_comp2$`X Dimension`)

WestVirginia_comp2$`Y Dimension` <- as.numeric(WestVirginia_comp2$`Y Dimension`)

WestVirginia_comp2 <- WestVirginia_comp2 %>% mutate('Possible_Taxon_Observation_Area' = `X Dimension` * `Y Dimension`)

```


Exporting full data table

```{r}

#As csv
write.csv(WestVirginia_comp2,"/home/shares/neon-inv/output_files/NCVS_WV_KPEACH_EB_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

WestVirginia_comp3 <- WestVirginia_comp2 %>%
  select(-'NEW_ExoticStatus', -'USDA_Growth_Form', -'USDA_Duration')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
WestVirginia_comp3 <- WestVirginia_comp3  %>% drop_na(Lat)

#As csv
write.csv(WestVirginia_comp3 ,"/home/shares/neon-inv/output_files/NCVS_WV_KPEACH_EB_reduced.csv", row.names = FALSE)

#write.csv(WestVirginia_comp3 ,"NCVS_WV_KPEACH_reduced.csv", row.names = FALSE)

```

