---
title: "West_Virginia_KPEACH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r}

library(readr)
library(readxl)
library(tidyverse)
library(knitr)
library(readxl)
library(openxlsx)
library(lubridate)

```

Reading in West Virginia Data (from a zip file delivered by email from Bob Peet)

```{r, message = FALSE}

# Paths to data
shared_data_dir <- "/home/shares/neon-inv"
data_raw <- file.path(shared_data_dir, "raw_VegBank_data")
data_output <- file.path(shared_data_dir, "output_files")

#New file path to shared folder
#Yikes so many errors
#Plot_species_list_WV <- read_excel(file.path(data_raw, "West_Virginia_cover_data.xlsx"),  sheet = "Plots-Species", guess_max = min(4000, n_max = NULL))

#Saved the sheets I needed as csv files
Plot_species_list_WV <- read_csv(file.path(data_raw, "WV_species_data.csv"))

Plot_data_WV <- read_csv(file.path(data_raw, "WV_plot_data.csv"))

#Import NEW Exotic Status Authority document
KPEACH_ExoticStatus_authority <- read_csv("/home/shares/neon-inv/output_files/KPEACH_ExoticStatus_authority_031621.csv")

```

Removing rows with NAs in critical columns

```{r}


Plot_data_WV <- Plot_data_WV %>% drop_na(`Survey Date`)

Plot_species_list_WV <- Plot_species_list_WV %>% drop_na(Stratum)



```

There are >1,000 semi-duplicate rows in the final merged dataset (All_VegBank_KPEACH_reduced.csv) so in this code chunk I am determining how many duplicates are in the data before I even really touch it.

More than 19,000 Duplicate row is in the data before I do anything to it( mostly due to the different stratum).

```{r}

duplicates1 <- Plot_species_list_WV %>% group_by(`Plot Code`, `Scientific Name`) %>% summarise(n_obs = n()) %>% filter(n_obs > 1) 

```

Selecting the columns I will need from the data to make a new dataset that will be compatible to merge with the PAINLES_25Oct2020 datasheet

```{r}



Plot_data_WV2 <- Plot_data_WV %>%
  select('Plot Code', 'Former Plot Code', 'Quad Code', 'Survey Date', 'Corrected Lat', 'Corrected Long', 'X Dimension', 'Y Dimension') %>%
  rename( 'UniqueID' = 'Plot Code','Year' = 'Survey Date', 'Lat' = 'Corrected Lat', 'Long' = 'Corrected Long' ) 

Plot_species_list_WV2 <- Plot_species_list_WV %>%
  select('Plot Code', 'Scientific Name', 'Stratum', 'Real Cover', 'Uncertain ID?', 'Plant Symbol') %>%
  rename( 'UniqueID' = 'Plot Code','bestname' = 'Scientific Name', 'PctCov' = 'Real Cover', 'SpCode' = 'Plant Symbol') %>% filter('Uncertain ID?' != TRUE) %>% select(-'Uncertain ID?')

WestVirginia <- left_join(Plot_data_WV2 , Plot_species_list_WV2, by = 'UniqueID')


```

```{r}

WestVirginia <- WestVirginia %>% drop_na(Year)

unique(WestVirginia$Year)

#Fixing date format
WestVirginia$Year <- lubridate::mdy(WestVirginia$Year)

#Reducing sampling date to just the year to match NEON data
WestVirginia <- WestVirginia %>% 
  mutate('Year' = substr(WestVirginia$Year, 1, 4))

unique(WestVirginia$Year)


#Adding a site column
WestVirginia <- WestVirginia %>% add_column(Dataset = 'NCVS_WV', .before = "UniqueID")  %>% add_column(Zone = 'L48', .before = "UniqueID")

WestVirginia$bestname <- gsub("\\bsp\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSP\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bL\\.\\b", " ", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname   <- str_replace_all(WestVirginia$bestname , "\\*|\\[|\\]", "")

WestVirginia$bestname   <- str_replace_all(WestVirginia$bestname , "\\*|\\{|\\}", "")

```

```{r}

unique(WestVirginia$Stratum)

WestVirginia$Stratum <- as.character(WestVirginia$Stratum)

WestVirginia <- WestVirginia %>% filter(Stratum != "T")

WestVirginia <- WestVirginia %>% filter(Stratum != "N")


```

Manually fixing names
```{r}

WestVirginia$bestname <- gsub("Carya alba (L.) Nutt. ex Ell.", "Carya tomentosa", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("CAAL27", "CATO6", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Polygonum caespitosum Blume var. longisetum (de Bruyn) A.N. Steward", "Polygonum cespitosum var. longisetum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("POCAL2", "POCEL", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Eupatorium purpureum L. var. purpureum", "Eutrochium purpureum var. purpureum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("EUPUP", "EUPUP4", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Dichanthelium dichotomum (L.) Gould ssp. dichotomum", "Dichanthelium dichotomum var. dichotomum", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("DIDIDWV", "DIDID", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("Acalypha gracilens Gray var. gracilens", "Acalypha gracilens", WestVirginia$bestname)

WestVirginia$SpCode <- gsub("ACGRGWV", "ACGR2", WestVirginia$SpCode)

WestVirginia$bestname <- gsub("\\bsnag\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSnag\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSnags\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bsnags\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSNAGS\\b", "", WestVirginia$bestname, ignore.case = FALSE)

WestVirginia$bestname <- gsub("\\bSNAG\\b", "", WestVirginia$bestname, ignore.case = FALSE)

#Filtering out anything listed as a snag (dead tree) 
WestVirginia <- WestVirginia %>% 
  filter(!bestname %in% c("snag", "Snag", "SNAG", "SNAGS", "snags", "Snags", "snag-", "Snags - ", "snag - "))


```



Table of species with no match in KPEACH_ExoticStatus_authority table

```{r}

#44 rows with no match after filtering out <.10 cover. These are all hybrids or observations with low confidence in species or genus identification (because multiple genera or species are listed)
no_match <- dplyr::anti_join(WestVirginia, KPEACH_ExoticStatus_authority, by="SpCode")


no_match <- no_match %>% drop_na(bestname)

to_be_matched <- no_match %>% 
    filter(!str_detect(bestname,  ".\\+|\\+."))

to_be_matched <- to_be_matched %>% 
    filter(bestname != "Lichen") %>%
    filter(bestname != "Liverwort") %>%
    filter(bestname != "Bryophyte") %>%
    filter(bestname != "Unknown") %>%
  filter(bestname != "snag") %>%
   filter(bestname != "snags") %>%
   filter(bestname != "SNAG") %>%
   filter(bestname != "Snag") %>%
 filter(!str_detect(bestname, "\\d"))

to_be_matched <- unique(to_be_matched)

```

Adding made up SpCodes for the hybrid rows

```{r}

#Adding new 'fake' species codes for the rows with no species codes (because they are hybrids or observer was uncertain of ID)

#Making a column of row numbers
no_match <- cbind(rownames(no_match), data.frame(no_match, row.names=NULL))

#Adding a - so each new species code will just have a 'NOMATCH' then a dash followed by sequential numbers
numbers <- paste0('-', no_match$`rownames(no_match)`)

#Final step
no_match <- no_match %>% 
mutate('SpCode' = paste0("NOMATCH_WV", numbers)) %>% 
select(-'rownames(no_match)') %>% 
rename('Former Plot Code' = 'Former.Plot.Code', 'Quad Code' = 'Quad.Code', 'X Dimension' = 'X.Dimension', 'Y Dimension' = 'Y.Dimension')



```

Adding species codes to rows with matches

```{r}


WestVirginia1 <- dplyr::inner_join(WestVirginia, KPEACH_ExoticStatus_authority, by="SpCode")

WestVirginia1 <- distinct(WestVirginia1)

WestVirginia1 <- WestVirginia1 %>%
  rename("bestname" = "bestname.y", "NCVS_original_scientific_name" = "bestname.x")

#Merging the df that DOES has correct species codes with the no_match table I made with the newly invented species codes

WestVirginia2 <- merge(WestVirginia1, no_match, all = TRUE)

WestVirginia3 <- WestVirginia2 %>%
  filter(is.na(Synonym.Symbol)) %>%
  select(-Synonym.Symbol)


```


```{r}

#Removing punctuation so I can extract only the first word from this column
WestVirginia3$NEON_GrowthForm <- gsub('[[:punct:]]', " ", WestVirginia3$NEON_GrowthForm)

#Extracting first word
WestVirginia3$NEON_GrowthForm <- stringr::word(WestVirginia3$NEON_GrowthForm, 1)

#Making new column
WestVirginia4 <- WestVirginia3  %>%
  mutate('NEW_GrowthForm' = ifelse(!is.na(NEON_GrowthForm), NEON_GrowthForm, USDA_Growth_Form))


```

Renaming USDA_Duration column to just 'Duration' 

```{r}

WestVirginia5 <- WestVirginia4 %>% rename('Duration' = 'USDA_Duration')


```

Making a possible Taxon Observation Area column
```{r}

WestVirginia5$`X Dimension` <- as.numeric(WestVirginia5$`X Dimension`)

WestVirginia5$`Y Dimension` <- as.numeric(WestVirginia5$`Y Dimension`)

WestVirginia5 <- WestVirginia5 %>% mutate('Possible_Taxon_Observation_Area' = `X Dimension` * `Y Dimension`)

```


Exporting full data table

```{r}

#As csv
write.csv(WestVirginia5,"/home/shares/neon-inv/output_files/NCVS_WV_KPEACH_ALL_COLS.csv", row.names = FALSE)

```

Exporting reduced data table

```{r}

WestVirginia6 <- WestVirginia5 %>%
  select(-'NEON_GrowthForm', -'USDA_Growth_Form', -'USDA_Exotic_Status')

#Just using this to see if any rows are dropped because there is no latitute and longitude. There really ought to be lat and long so just trying to catch that here
WestVirginia6 <- WestVirginia6  %>% drop_na(Lat)

#As csv
write.csv(WestVirginia6 ,"/home/shares/neon-inv/output_files/NCVS_WV_KPEACH_reduced.csv", row.names = FALSE)

write.csv(WestVirginia6 ,"NCVS_WV_KPEACH_reduced.csv", row.names = FALSE)

```
